### 2024-02-12 article_jos.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
]{article}

\usepackage{amsmath,amssymb}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[top=2cm,bottom=2cm,left = 2.5cm,right = 2.5cm]{geometry}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{241,243,245}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.40,0.45,0.13}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\ExtensionTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.28,0.35,0.67}{#1}}
\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.00,0.46,0.62}{#1}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\NormalTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.07,0.07,0.07}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage{stmaryrd}
\usepackage{multicol}
\usepackage{graphicx}
\usepackage{animate, dsfont, here, xspace}
%\usepackage{tikz}
\usepackage{tikz,pgfplots}
 \pgfplotsset{compat=1.17}
%\includepdf[fitpaper=true, pages=-]{img/pdg.pdf}

\DeclareMathOperator{\e}{e}
\DeclareMathOperator{\Determinant}{det}
\renewcommand{\P}{\mathds{P}} %Apparement \P existe déjà ?
\newcommand\N{\mathds{N}}
\newcommand\Norm{\mathcal{N}}
\newcommand\R{\mathds{R}}
\newcommand\Z{\mathds{Z}}
\newcommand{\determinant}[1]{\Determinant\left(#1\right)}
%\newcommand\C{\mathds{C}}
%\newcommand\Z{\mathds{Z}}

\newcommand\1{\mathds{1}}
\newcommand{\E}[2][]{{\mathds{E}}_{#1}
  \def\temp{#2}\ifx\temp\empty
  \else
    \left[#2\right]
  \fi
}
\newcommand{\V}[2][]{{\mathds{V}}_{#1}
  \def\temp{#2}\ifx\temp\empty
  \else
    \left[#2\right]
  \fi
}
\newcommand\ud{\,\mathrm{d}}
\newcommand{\ps}[2]{\left\langle #1 \,,\, #2 \right\rangle}

% blocks
\usepackage{environ}
\usepackage[tikz]{bclogo}

\tikzstyle{titlestyle} =[draw=black!80,fill=black!20, text=black,
 right=10pt, rounded corners]
\mdfdefinestyle{symmaryboxstyle}{
	linecolor=black!80, backgroundcolor = black!5,
	skipabove=\baselineskip, innertopmargin=\baselineskip,
	innerbottommargin=\baselineskip,
	userdefinedwidth=\textwidth,
	middlelinewidth=1.2pt, roundcorner=5pt,
	skipabove={\dimexpr0.5\baselineskip+\topskip\relax},
	frametitleaboveskip=\dimexpr-\ht\strutbox\relax,
	innerlinewidth=0pt,
}
\newcounter{summarybox}
\NewEnviron{summary_box}[2][true]{%
\refstepcounter{summarybox} % on incrémente le compteur
\begin{mdframed}[style=symmaryboxstyle,
nobreak=#1,
frametitle={%
      \tikz[baseline=(current bounding box.east),outer sep=0pt]
      \node[titlestyle,anchor=east]
    {Encadré \thesummarybox{} --- #2};}
]
\vspace{-0.5em}
\BODY
\end{mdframed}
}

\usepackage{amsthm}
%\theoremstyle{remark}
\newtheorem*{remarque}{Remarque}

\usepackage{mathrsfs}
\usepackage{fontawesome5}

%\usepackage[style=authoryear,%,uniquename=false, uniquelist=false,
%maxbibnames=2]{biblatex-chicago}
% \usepackage[options]{natbib}
% \bibliographystyle{chicago}

%\DefineBibliographyStrings{english}{andothers={et\addabbrvspace alii}}
\usepackage[authordate16,backend=biber,maxcitenames=2]{biblatex-chicago}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{fontawesome5}{}{\usepackage{fontawesome5}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{biblatex}
\addbibresource{biblio.bib}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Improving real-time trend estimates using local parametrisation of local polynomial regression filters},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Improving real-time trend estimates using local parametrisation
of local polynomial regression filters}
\author{}
\date{}

\begin{document}
\maketitle
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, interior hidden, sharp corners, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, breakable, frame hidden]}{\end{tcolorbox}}\fi

\setstretch{2}
\hypertarget{abstract}{%
\section*{Abstract}\label{abstract}}
\addcontentsline{toc}{section}{Abstract}

This paper studies and compares real-time estimates of the trend cycle
component with moving averages based on local polynomial regression. In
particular, this theory allows to reproduce Henderson's symmetric and
Musgrave's asymmetric filters. This paper describes how they can be
extended to include a timeliness criterion to minimise the phase shift.
While asymmetric filters are generally parameterised globally, which can
be suboptimal around turning points, this paper proposes a procedure for
parameterising them locally. An empirical comparison, based on simulated
and real data, shows that modelling polynomial trends that are too
complex introduces more revisions without reducing the phase shift, and
that local parameterisation reduces the delay in detecting turning
points and reduces revisions. The paper also shows how these results can
be easily reproduced using the R package \texttt{rjd3filters} dedicated
to the manipulation of moving averages.

Keywords: time series, trend-cycle, seasonal adjustment, turning points,
R statistical software.

JEL Classification: E32, E37.

\hypertarget{introduction}{%
\section*{Introduction}\label{introduction}}
\addcontentsline{toc}{section}{Introduction}

Analysis of the economic cycle, and in particular the early detection of
turning points in a series, is a major topic in the analysis of economic
outlook. To this end, economic indicators are generally seasonally
adjusted. However, in order to improve their interpretability, it may be
necessary to perform additional smoothing in order to reduce noise, and
thus analyse the trend-cycle component. By construction, trend-cycle
extraction methods are closely related to seasonal adjustment methods,
since they are generally applied to seasonally adjusted series.

A moving average, or linear filter, is a statistical method that
consists in applying a rolling weighted mean to a times series: for each
date \(t\) it computes a weighted mean of \(p\) past points and \(q\)
future points where \(p,q\geq0\) depends on the moving average. Moving
averages are ubiquitous in business cycle extraction and seasonal
adjustment methods. For example, the X-13ARIMA seasonal adjustment
method uses several moving averages to estimate the main components of a
time series. Symmetric filters are applied to the center of the series,
but when it comes to estimate the most recent points, all of these
methods must rely on asymmetric filters. For trend-cycle extraction, the
most popular symmetric filter is the \textcite{henderson1916note} moving
average, which is used in the X-13ARIMA seasonal adjustment algorithm.

However, for real-time estimates, due to the lack of future
observations, all these methods must rely on asymmetric filters to
estimate the most recent points. For example, for trend-cycle
extraction, X-13ARIMA uses Henderson's symmetric filter and
\textcite{musgrave1964set}'s asymmetric filters on an extended series
using an ARIMA model. Since the predicted values are linear combinations
of past values, this amounts to applying asymmetric averages at the end
of the series.

If the classic asymmetric moving averages have good properties regarding
the future revisions induced by the process (see for example
\textcite{pierce1980SA}), they create, by construction, phase shifts
that impact the real-time estimation of turning points, introducing time
delay in the detection.

Using local polynomial regression, \textcite{proietti2008} developed
general class of symmetric and asymmetric moving averages. In
particular, it makes possible to reproduce Henderson's symmetric filter
and Musgrave's asymmetric filters. However, these methods have two
drawbacks. Firstly, the phase shift (i.e.~the delay in detecting turning
points) is not directly controlled. Secondly, asymmetric filters depend
on a parameter that is generally estimated over the whole series using
assumptions that may be wrong locally. The aim of this study is to
propose extensions to the \textcite{proietti2008} class of filters in
order to take these shortcomings into account. The extensions have been
implemented in the \faIcon{r-project} \texttt{rjd3filters} package
(available at \url{https://github.com/rjdemetra/rjd3filters}) and are
therefore easily reusable.

In Section~\ref{sec-propMM}, we describe the general properties of
moving averages and the associated quality criteria. This allows us to
understand the foundations behind the construction of moving averages
from local polynomial regressions, as well as those behind the two
extensions proposed in this article (Section~\ref{sec-nonparamreg}). In
a final section (Section~\ref{sec-comparison}), all these methods are
compared empirically on simulated and real series.

\hypertarget{sec-propMM}{%
\section{Some properties on moving averages}\label{sec-propMM}}

Lots of papers describe the definition and the properties of moving
averages and linear filters, see for example \textcite{ch12HBSA}. In
this section we summarize some of the main results to understand the
next sections.

Let \(p\) et \(f\) two integers, a moving average
\(M_{\boldsymbol\theta}\) or \(M\) is defined by a set of coefficients
\(\boldsymbol\theta=\begin{pmatrix}\theta_{-p}\\\vdots\\\theta_{f}\end{pmatrix}\)
such as for all time series \(X_t\): \[
M_{\boldsymbol\theta}(X_t)=\sum_{k=-p}^{+f}\theta_kX_{t+k}.
\]

\begin{itemize}
\item
  \(p+f+1\) is called the \emph{moving average order}.
\item
  When \(p=f\) the moving average is said to be \emph{centered}. If we
  also have \(\forall k:\:\theta_{-k} = \theta_k\), the moving average
  \(M_{\boldsymbol\theta}\) is said to be \emph{symmetric}. In this
  case, the quantity \(h=p=f\) is called the \emph{bandwidth}.
\end{itemize}

\hypertarget{subsec:gain-deph}{%
\subsection{Gain and phase shift functions}\label{subsec:gain-deph}}

To interpret the notions of gain and phase shift, it is useful to
illustrate the effects of moving averages on harmonic series
\(X_t=\e^{-i\omega t}\) with \(\omega\in[0,\pi]\). The moving average
\(M_{\boldsymbol\theta}\) transforms \(X_t\) into: \[
Y_t = M_{\boldsymbol\theta}X_t = \sum_{k=-p}^{+f} \theta_k \e^{-i \omega (t+k)}
= \left(\sum_{k=-p}^{+f} \theta_k \e^{-i \omega k}\right)\cdot X_t.
\] The function
\(\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k e^{-i \omega k}\)
is called the \emph{transfer function} or \emph{frequency response
function}. The frequency response function can equivalently be defined
as
\(\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k e^{i \omega k}\)
ou
\(\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k e^{2\pi i \omega k}\).
It can be rewritten as: \[
\Gamma_{\boldsymbol\theta}(\omega) = \rho_{\boldsymbol\theta}(\omega)\e^{i\varphi_{\boldsymbol\theta}(\omega)},
\] where
\(\rho_{\boldsymbol\theta}(\omega)=G_{\boldsymbol\theta}(\omega)=\lvert\Gamma_{\boldsymbol\theta}(\omega)\rvert\)
is the \emph{gain} or \emph{amplitude} function and
\(\varphi_{\boldsymbol\theta}(\omega)\) is the \emph{phase shift} or
\emph{time shift} function. This function is sometimes represented as
\(\phi_{\boldsymbol\theta}(\omega)=\frac{\varphi_{\boldsymbol\theta}(\omega)}{\omega}\)
to mesure the phase shift in number of periods. For all symmetric moving
average we have
\(\varphi_{\boldsymbol\theta}(\omega)\equiv 0 \;(modulo\;{\pi})\).

To sum up, applying a moving average to a harmonic times series
(\(X_t=\e^{-i\omega t}\)) affects it in in two different ways:

\begin{itemize}
\item
  by multiplying it by an amplitude coefficient
  \(\rho_{\boldsymbol\theta}\left(\omega\right)\) (gain);
\item
  by ``shifting'' it in time by
  \(\varphi_{\boldsymbol\theta}(\omega)/\omega\) which directly affects
  the detection of turning points. When
  \(\varphi_{\boldsymbol\theta}(\omega)/\omega<0\) the time shift is
  negative: the turning point is detected with delay.
\end{itemize}

Fourier decomposition allows to analyze any time series as a sum of
harmonic series, and each component (trend, cycle, seasonal, irregular)
is associated with a set of frequencies. Noting \(\omega = 2\pi/p\), the
harmonic series of frequency \(\omega\) represents a series that repeats
itself every \(p\) periods. For example, for a monthly series (12
observations per year), seasonal movements are those that repeat every
year: they are therefore associated with frequencies \(2\pi/12\) (annual
periodicity), \(2\pi/12\times 2=2\pi/6,\dots,2\pi/12\times 5\). In this
paper, we consider that the trend-cycle is associated with frequencies
in the interval \([0, 2\pi/12[\), i.e.~movements recurring at least
every 12 months. Even if different frequencies are sometimes used
(e.g.~\([0, 2\pi/36]\) to consider only cycles of at least 36 months),
this has no impact on the results of the study. The other frequencies,
\([2\pi/12, \pi]\) are associated with the irregular component
(undesirable oscillations).

Figure~\ref{fig-graphsmusgrave} shows the gain and phase shift function
for the asymmetric Musgrave filter (see Section~\ref{sec-lppasymf})
often used for real-time trend-cycle estimation (i.e.~when no point in
the future is known). The gain function is greater than 1 on the
frequencies associated with the trend-cycle (\([0, 2\pi/12]\)): this
means that the trend-cycle is well preserved and that short cycles of 1
to 2 years (\([2\pi/24,2\pi/12]\)) are even amplified. However, cycles
of 8 to 12 months (\([2\pi/12, 2\pi/8]\)), considered undesirable
because associated with the irregular, are also amplified: this can lead
to the detection of false turning points. At other frequencies, the gain
function is less than 1, but always positive: this means that the series
smoothed by this moving average will always contain noise, even if it is
attenuated. Analysis of the phase shift shows that the shorter the
cycles, the greater the phase shift: this means that on series smoothed
by this moving average, turning points could be detected at the wrong
date.

\begin{figure}[H]

\caption{\label{fig-graphsmusgrave}Coefficients, gain and phase shift
function for the Musgrave filter in real time with \(I/C=3.5\).}

{\centering \includegraphics[width=1\textwidth,height=\textheight]{img/filters_used/musgrave.pdf}

}

\end{figure}

\hypertarget{desirable-properties-of-a-moving-average}{%
\subsection{Desirable properties of a moving
average}\label{desirable-properties-of-a-moving-average}}

To decompose a time series into a seasonal component, a trend-cycle and
the irregular, the X-11 decomposition algorithm (used in X-13ARIMA) uses
a succession of moving averages, all with specific constraints.

In our case, we assume that our initial series \(y_t\) is seasonally
adjusted and can be written as the sum of a trend-cycle, \(TC_t\), and
an irregular component, \(I_t\): \[
y_t=TC_t+I_t.
\] The aim will be to build moving averages that best preserve the
trend-cycle (\(M_{\boldsymbol\theta} (TC_t)\simeq TC_t\)) and minimise
noise (\(M_{\boldsymbol\theta} (I_t)\simeq 0\)).

\hypertarget{trend-preservation}{%
\subsubsection{Trend preservation}\label{trend-preservation}}

Trend-cycles are generally modelled by local polynomial trends (see
Section~\ref{sec-nonparamreg}). In order to best preserve the
trend-cycles, we try to have moving averages that preserve the
polynomial trends. A moving average \(M_{\boldsymbol\theta}\) preserves
a function of time \(f(t)\) if
\(\forall t:M_{\boldsymbol\theta} f(t)=f(t)\).

We have the following properties for the moving average
\(M_{\boldsymbol\theta}\):

\begin{itemize}
\item
  To preserve a constant trend \(X_t=a\) we need: \[
  \forall t:M_{\boldsymbol\theta}(X_t)=\sum_{k=-p}^{+f}\theta_kX_{t+k}=\sum_{k=-p}^{+f}\theta_ka=a\sum_{k=-p}^{+f}\theta_k=a.
  \] The sum of the coefficients of the moving average
  \(\sum_{k=-p}^{+f}\theta_k\) must then be equal to \(1\).
\item
  To preserve linear trends \(X_t=at+b\) we need: \[
  \forall t:\:M_{\boldsymbol\theta}(X_t)=\sum_{k=-p}^{+f}\theta_kX_{t+k}=\sum_{k=-p}^{+f}\theta_k[a(t+k)+b]=a\sum_{k=-p}^{+f}k\theta_k+(at+b)\sum_{k=-p}^{+f}\theta_k=at+b.
  \] Which is equivalent to: \[
  \sum_{k=-p}^{+f}\theta_k=1
  \quad\text{et}\quad
  \sum_{k=-p}^{+f}k\theta_k=0.
  \]
\item
  In general, \(M_{\boldsymbol\theta}\) preserves polynomials of degree
  \(d\) if and only if: \[
  \sum_{k=-p}^{+f}\theta_k=1
   \text{ and }
  \forall j \in \left\llbracket 1,d\right\rrbracket:\:
  \sum_{k=-p}^{+f}k^j\theta_k=0.
  \]
\item
  If \(M_{\boldsymbol\theta}\) is symmetric (\(p=f\) and
  \(\theta_{-k} = \theta_k\)) and preserves polynomials of degree \(2d\)
  then it also preserves polynomials of degree \(2d+1\).
\end{itemize}

\hypertarget{noise-reduction}{%
\subsubsection{Noise reduction}\label{noise-reduction}}

All time series are affected by noise, which can blur the extraction of
trends and cycles. This is why we try to reduce this noise (by reducing
its variance) while preserving the relevant trends. The sum of the
squares of the coefficients \(\sum_{k=-p}^{+f}\theta_k^2\) is the
\emph{reduction of variance} ratio.

Indeed, let \(\{\varepsilon_t\}\) a sequence of independent random
variables with \(\E{\varepsilon_t}=0\), \(\V{\varepsilon_t}=\sigma^2\):
\[
\V{M_{\boldsymbol\theta}\varepsilon_t}=\V{\sum_{k=-p}^{+f} \theta_k \varepsilon_{t+k}}
= \sum_{k=-p}^{+f} \theta_k^2 \V{\varepsilon_{t+k}}=
\sigma^2\sum_{k=-p}^{+f} \theta_k^2.
\]

\hypertarget{sec-mmasym}{%
\subsection{Real-time estimation and asymmetric moving
average}\label{sec-mmasym}}

\hypertarget{sec-mmetprev}{%
\subsubsection{Asymmetric moving averages and
forecasts}\label{sec-mmetprev}}

For symmetric filters, the phase shift function is equal to zero (modulo
\(\pi\)). So there is no delay in detecting turning points. However,
they cannot be used in the beginning and in the end of the time series
because no past or future value can be used. Thus, for real-time
estimation, it is needed to build asymmetric moving average that
approximate the symmetric moving average.

Several approaches can be used for real-time estimation:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Use asymmetric moving averages to take account of the lack of
  available data;
\item
  Apply symmetrical filters to extended forecast series. This method
  seems to date back to \textcite{deforest1877adjustment}, which also
  suggests modelling a polynomial trend of degree three or less at the
  end of the period:\\
  ``\emph{As the first \(m\) and last \(m\) terms of the series cannot
  be reached directly by the formula, the series should be graphically
  extended by m terms at both ends, first plotting the observations on
  paper as ordinates, and then extending the curve along what seems to
  be its probable course, and measuring the ordinates of the extended
  portions.} \emph{It is not necessary that this extension should
  coincide with what would be the true course of the curve in those
  parts. } \emph{The important point is that the m terms thus added,
  taken together with the \(m+1\) adjacent given terms, should follow a
  curve whose form is approximately algebraic and of a degree not higher
  than the third.}''\\
  This is also the approach used in the X-13ARIMA seasonal adjustment
  method, which extends which extends the series by one year with an
  ARIMA model.
\end{enumerate}

Ultimately, the second method is equivalent to using asymmetric moving
averages, since forecasts are linear combinations of the past.

Conversely, the implicit forecasts of an asymmetric moving average can
be deduced from a reference symmetric moving average. This allows to
judge the quality of real-time estimates of the trend-cycle and to
anticipate future revisions when forecasts are far from expected
evolutions.\\
Let us denote \(v=(v_{-h},\dots, v_{h})\) the reference symmetrical
moving average and \(w^0,\dots w^{h-1}\) a sequence of asymmetrical
moving averages, of order \(h+1\) to \(2h\) used to estimate the last
\(h\) points with, for convention, \(w_t^q=0\) for \(t>q\). This means
that \(w^0=(w_{-h}^0,\dots, w_{0}^0)\) is used for estimation in real
time (when no points are known in the future),
\(w^1=(w_{-h}^1,\dots, w_{1}^1)\) for estimation of the penultimate
point (when only one point is known in the future), and so on. Let also
note \(y_{-h},\dots,y_{0}\) the observed series studied and
\(y_{1}^*,\dots y_h^*\) the implicit forecast induced by
\(w^0,\dots w^{h-1}\). This means that for all \(q\) we have: \[
\forall q, \quad \underbrace{\sum_{i=-h}^0 v_iy_i + \sum_{i=1}^h v_iy_i^*}_{\text{smoothing by }v\text{ of the extended series}}
=\underbrace{\sum_{i=-h}^0 w_i^qy_i + \sum_{i=1}^h w_i^qy_i^*}_{\text{smoothing by }w^q\text{ of the extended series}}.
\] This is equivalent to: \[
\forall q, \quad \sum_{i=1}^h (v_i- w_i^q) y_i^*
=\sum_{i=-h}^0 (w_i^q-v_i)y_i.
\] To sum up, in matrix form, this is equivalent to solving: \[
\scriptstyle
\begin{pmatrix}
  v_1 & v_2 & \cdots & v_h \\
  v_1 - w_1^1 & v_2 & \cdots & v_h \\
  \vdots & \vdots & \cdots & \vdots \\
   v_1 - w_1^{h-1} & v_2-w_2^{h-1} & \cdots & v_h
\end{pmatrix}
\begin{pmatrix}y_1^* \\ \vdots \\ y_h^*\end{pmatrix}=
\begin{pmatrix}
  w_{-h}^0 - v_{-h} & w_{-(h-1)}^0 - v_{-(h-1)} & \cdots & w_{0}^0 - v_{0} \\
  w_{-h}^1 - v_{-h} & w_{-(h-1)}^1 - v_{-(h-1)} & \cdots & w_{0}^1 - v_{0} \\
  \vdots & \vdots & \cdots & \vdots \\
  w_{-h}^{h-1} - v_{-h} & w_{-(h-1)}^{h-1} - v_{-(h-1)} & \cdots & w_{0}^{h-1} - v_{0}
\end{pmatrix}
\begin{pmatrix}y_{-h} \\ \vdots \\ y_0\end{pmatrix}.
\] This is implemented in the \texttt{rjd3filters::implicit\_forecast()}
function.

As highlighted by \textcite{wildischis2004}, extending the series
through forecasting with an ARIMA model is equivalent to calculating
asymmetric filters whose coefficients are optimised in relation to the
one-step ahead forecast. In other words, the aim is to minimise the
revisions between the first and last estimates (with the symmetric
filter). However, the phase shift induced by the asymmetric filters is
not controlled: we might prefer to have faster detection of turning
points and a larger revision rather than just minimising the revisions
between the first and last estimates. Furthermore, since the
coefficients of the symmetric filter (and therefore the weight
associated with distant forecasts) decrease slowly, we should also be
interested in the performance of multi-step ahead forecasting. This is
why it may be necessary to define alternative criteria for judging the
quality of asymmetric moving averages.

\hypertarget{subsec:crit-qual}{%
\subsubsection{Quality indicators for asymmetric moving
averages}\label{subsec:crit-qual}}

For asymmetric filters, most of the criteria come from those defined by
\textcite{ch15HBSA} and \textcite{trilemmaWMR2019} to build asymmetric
filters. They are summarised in Table~\ref{tab-QC} and can be calculated
using the \texttt{rjd3filters::diagnostic\_matrix()} function.

\textcite{ch15HBSA} propose a general approach to derive linear filters,
based on an optimisation problem involving three criteria:
\emph{Fidelity} (\(F_g\)), \emph{Smoothness} (\(S_g\)) and
\emph{Timeliness} (\(T_g\)). Fidelity can be directly linked to the
reduction in variance created by the filter, and timeliness to the
notion of phase shift, which again we want to be low. This method can be
used in\faIcon{r-project} with the function
\texttt{rjd3filters::fst\_filter()}.

\textcite{trilemmaWMR2019} propose an approach based on the
decomposition of the mean squared error between the symmetric and the
asymmetric filter in four quantities: \emph{Accuracy} (\(A_w\)),
\emph{Timeliness} (\(T_w\)), \emph{Smoothness} (\(S_w\)) and
\emph{Residual} (\(R_w\)). A simplified version of this method can be
used in \faIcon{r-project} with the function
\texttt{rjd3filters::dfa\_filter()}.

\begin{table}[!h]
\caption{\label{tab-QC} Quality criteria for asymmetric filters ($q=0,1,2$) computed by local polynomials using the Henderson kernel with $h=6$ and $R=3,5$.}
{\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{ccc}
\toprule
Code & Description & Formula\\
\midrule
$b_c$ & Constant bias & $\sum_{k=-p}^{+f}\theta_{k}-1$\\
$b_l$ & Linear bias & $\sum_{k=-p}^{+f}k\theta_{k}$\\
$b_q$ & Quadratic bias & $\sum_{k=-p}^{+f}k^{2}\theta_{k}$\\
$F_g$ & Variance reduction / Fidelity (Guggemos) & $\sum_{k=-p}^{+f}\theta_{k}^{2}$\\
$S_g$ & Smoothness (Guggemos) & $\sum_{j}(\nabla^{3}\theta_{j})^{2}$\\
\addlinespace
$T_g$ & Timeliness (Guggemos) & $\int_{0}^{\omega_1}\rho_{\boldsymbol\theta}(\omega)\sin(\varphi_{\boldsymbol\theta}(\omega))^{2}\ud\omega$\\
$A_w$ & Accuracy (Wildi) & $2\int_0^{\omega_1}\left(\rho_{s}(\omega)-\rho_{\boldsymbol\theta}(\omega)\right)^{2}h(\omega)\ud\omega$\\
$T_w$ & Timeliness (Wildi) & $8\int_0^{\omega_1} \rho_{s}(\omega)\rho_{\boldsymbol\theta}(\omega)\sin^{2}\left(\frac{\varphi_s(\omega)-\varphi_{\boldsymbol\theta}(\omega)}{2}\right)h(\omega)\ud\omega$\\
$S_w$ & Smoothness (Wildi) & $2\int_{\omega_1}^{\pi}\left(\rho_{s}(\omega)-\rho_{\boldsymbol\theta}(\omega)\right)^{2}h(\omega)\ud\omega$\\
$R_w$ & Residual (Wildi) & $8\int_{\omega_1}^{\pi} \rho_{s}(\omega)\rho_{\boldsymbol\theta}(\omega)\sin^{2}\left(\frac{\varphi_s(\omega)-\varphi_{\boldsymbol\theta}(\omega)}{2}\right)h(\omega)\ud\omega$\\
\bottomrule
\end{tabular}}}

\textit{Note: $X_g$ criteria from Grun-Rehomme et al (2018) and $X_w$ criteria from Wildi and McElroy (2019).}

\textit{$\rho_s$ and $\varphi_s$ represent the gain and phase shift function of the symmetric Henderson filter.}

\textit{$h$ is the spectral density of the input series, considered to be that of white noise ($h_{WN}(x)=1$) or a random walk ($h_{RW}(x)=\frac{1}{2(1-\cos(x))}$).}
\end{table}

\hypertarget{sec-nonparamreg}{%
\section{Non-parametric regression and local polynomial
regression}\label{sec-nonparamreg}}

Many trend-cycle extraction methods are based on non-parametric
regressions, which are particularly flexible because they do not assume
any predetermined dependency in the predictors. In practice, local
regressions can be used. More specifically, consider a set of points
\((x_i,y_i)_{1\leq i\leq n}\). Non-parametric regression involves
assuming that there exists a function \(\mu\), to be estimated, such
that \(y_i=\mu(x_i)+\varepsilon_i\) with \(\varepsilon_i\) an error
term. According to Taylor's theorem, for any point \(x_0\), if \(\mu\)
is differentiable \(d\) times, then:
\begin{equation}\protect\hypertarget{eq-taylor}{}{
\forall x \::\:\mu(x) = \mu(x_0) + \mu'(x_0)(x-x_0)+\dots +
\frac{\mu^{(d)}(x_0)}{d!}(x-x_0)^d+R_d(x),
}\label{eq-taylor}\end{equation} where \(R_d\) is a negligible residual
term in the neighbourhood of \(x_0\). In a neighbourhood \(h(x_0)\)
around \(x_0\), \(\mu\) can be approximated by a polynomial of degree
\(d\). The quantity \(h(x_0)\) is called the \emph{bandwidth}. If
\(\varepsilon_i\) is white noise, then we can estimate by least squares
\(\mu(x_0)\) using the observations which are in
\(\left[x_0-h(x_0),x_0+h(x_0)\right]\).

In practice, this means assuming that the trend is locally polynomial.
Various estimation methods can be used to derive symmetrical and
asymmetrical moving averages.\\
\textcite{GrayThomson1996} propose a complete statistical framework
which makes it possible, in particular, to model the error in
approximating the trend by local polynomials. However, as the
specification of this error is generally complex, simpler models may be
preferred, such as that of \textcite{proietti2008}.\\
Finally, \textcite{dagumbianconcini2008} propose a similar modelling of
the trend-cycle but using the theory of Hilbert spaces with reproducing
kernels for estimation, which has the particular advantage of
facilitating the calculation of different moving averages at different
time frequencies. This method can be used in\faIcon{r-project} with the
function \texttt{rjd3filters::rkhs\_filter()}.

\hypertarget{symmetric-moving-averages-and-local-polynomial-regression}{%
\subsection{Symmetric moving averages and local polynomial
regression}\label{symmetric-moving-averages-and-local-polynomial-regression}}

Using \textcite{proietti2008}'s notation, we assume that our time series
\(y_t\) can be decomposed into \[
y_t=\mu_t+\varepsilon_t,
\] where \(\mu_t\) is the trend and
\(\varepsilon_{t}\overset{i.i.d}{\sim}\mathcal{N}(0,\sigma^{2})\) is the
noise (the series is therefore seasonally adjusted). Within a
neighbourhood \(h\) of \(t\), the local trend \(\mu_t\) is approximated
by a polynomial of degree \(d\), such that \(\mu_t\simeq m_{t}\) with:
\[
\forall j\in\left\llbracket -h,h\right\rrbracket:\:
y_{t+j}=m_{t+j}+\varepsilon_{t+j},\quad m_{t+j}=\sum_{i=0}^{d}\beta_{i}j^{i}.
\] The problem of trend extraction is equivalent to estimating
\(m_t=\beta_0\) (the constant in the previous formula).

In matrix notation: \[
\underbrace{\begin{pmatrix}y_{t-h}\\
y_{t-(h-1)}\\
\vdots\\
y_{t}\\
\vdots\\
y_{t+(h-1)}\\
y_{t+h}
\end{pmatrix}}_{\boldsymbol y}=\underbrace{\begin{pmatrix}1 & -h & h^{2} & \cdots & (-h)^{d}\\
1 & -(h-1) & (h-1)^{2} & \cdots & (-(h-1))^{d}\\
\vdots & \vdots & \vdots & \cdots & \vdots\\
1 & 0 & 0 & \cdots & 0\\
\vdots & \vdots & \vdots & \cdots & \vdots\\
1 & h-1 & (h-1)^{2} & \cdots & (h-1)^{d}\\
1 & h & h^{2} & \cdots & h^{d}
\end{pmatrix}}_{\boldsymbol X}\underbrace{\begin{pmatrix}\beta_{0}\\
\beta_{1}\\
\vdots\\
\vdots\\
\vdots\\
\vdots\\
\beta_{d}
\end{pmatrix}}_{\boldsymbol \beta}+\underbrace{\begin{pmatrix}\varepsilon_{t-h}\\
\varepsilon_{t-(h-1)}\\
\vdots\\
\varepsilon_{t}\\
\vdots\\
\varepsilon_{t+(h-1)}\\
\varepsilon_{t+h}
\end{pmatrix}}_{\boldsymbol \varepsilon}.
\]

To estimate \(\boldsymbol \beta\) we need \(2h+1\geq d+1\) and the
estimation is made by weighted least squares (WLS), which is equivalent
to minimising the following objective function: \[
S(\hat{\beta}_{0},\dots,\hat{\beta}_{d})=\sum_{j=-h}^{h}\kappa_{j}(y_{t+j}-\hat{\beta}_{0}-\hat{\beta}_{1}j-\dots-\hat{\beta}_{d}j^{d})^{2}.
\] where \(\kappa_j\) is a set of weights called \emph{kernel}. We have
\(\kappa_j\geq 0:\kappa_{-j}=\kappa_j\), and with
\(\boldsymbol K=diag(\kappa_{-h},\dots,\kappa_{h})\), the estimate of
\(\boldsymbol \beta\) can be written as
\(\hat{\boldsymbol\beta}=(\boldsymbol X'\boldsymbol K\boldsymbol X)^{-1}\boldsymbol X'\boldsymbol K\boldsymbol y.\)
With \(\boldsymbol e_1=\begin{pmatrix}1 \\0 \\\vdots\\0 \end{pmatrix}\),
the estimate of the trend is:
\begin{equation}\protect\hypertarget{eq-mmsym}{}{
\hat{m}_{t}=\boldsymbol e_{1}\hat{\boldsymbol \beta}=\boldsymbol \theta'\boldsymbol y=\sum_{j=-h}^{h}\theta_{j}y_{t-j}\text{ with }\boldsymbol \theta=\boldsymbol K\boldsymbol X(\boldsymbol X'\boldsymbol K\boldsymbol X)^{-1}\boldsymbol e_{1}.
}\label{eq-mmsym}\end{equation} To conclude, the estimate of the trend
\(\hat{m}_{t}\) can be obtained applying the symmetric filter
\(\boldsymbol \theta\) to \(y_t\) (\(\boldsymbol \theta\) is symmetric
due to the symmetry of the kernel weights \(\kappa_j\)). Moreover,
\(\boldsymbol X'\boldsymbol \theta=\boldsymbol e_{1}\) so: \[
\sum_{j=-h}^{h}\theta_{j}=1,\quad\forall r\in\left\llbracket 1,d\right\rrbracket:\sum_{j=-h}^{h}j^{r}\theta_{j}=0.
\] Hence, the filter \(\boldsymbol w\) preserve deterministic polynomial
of order \(d\).

Regarding parameter selection, the general consensus is that the choice
between different kernels is not crucial. See, for example,
\textcite{cleveland1996smoothing} or \textcite{Loader1999}. The only
desired constraints on the kernel are that it assigns greater weight to
the central estimation (\(\kappa_0\)) and decreases towards 0 as it
moves away from the central estimation. The uniform kernel should
therefore be avoided. It is then preferable to focus on two other
parameters:

\begin{itemize}
\item
  the degree of the polynomial, denoted by \(d\): if it is too small,
  there is a risk of biased estimates of the trend-cycle, and if it is
  too large, there is a risk of excessive variance in the estimates (due
  to over-adjustment);
\item
  the number of neighbors \(2h+1\) (or the window \(h\)): if it is too
  small, then there will be too little data for the estimates (resulting
  in high variance in the estimates), and if it is too large, the
  polynomial approximation will probably be wrong, leading to biased
  estimates.
\end{itemize}

In this paper we will use the Henderson kernel: \[
\kappa_{j}=\left[1-\frac{j^2}{(h+1)^2}\right]
\left[1-\frac{j^2}{(h+2)^2}\right]
\left[1-\frac{j^2}{(h+3)^2}\right].
\] However, several other kernels are available in \texttt{rjd3filters}.

\hypertarget{sec-lppasymf}{%
\subsection{Asymmetric moving averages and local polynomial
regression}\label{sec-lppasymf}}

As mentioned in Section~\ref{sec-mmetprev}, for real-time estimation,
several approaches can be used:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Apply symmetric filters to the series extended by forecasting
  \(\hat{y}_{n+l\mid n},l\in\left\llbracket 1,h\right\rrbracket\).
\item
  Build an asymmetric filter by local polynomial approximation on the
  available observations (\(y_{t}\) for
  \(t\in\left\llbracket n-h,n\right\rrbracket\)).
\item
  Build asymmetric filters that minimise the mean squared error of
  revision under polynomial trend reproduction constraints.
\end{enumerate}

\textcite{proietti2008} show that the first two approaches are
equivalent when forecasts are made by polynomial extrapolation of degree
\(d\). They are also equivalent to the third approach under the same
constraints as those of the symmetric filter. The third method is called
\emph{direct asymmetric filter} (DAF). This method is used for real-time
estimation in the STL seasonal adjustment method (\emph{Seasonal-Trend
decomposition based on Loess}, see \textcite{cleveland90}). Although DAF
filter estimates are unbiased, this is at the cost of greater variance
in the estimates.

To solve the problem of the variance of real-time filter estimates,
\textcite{proietti2008} propose a general method for constructing
asymmetric filters that allows a bias-variance trade-off. This is a
generalisation of \textcite{musgrave1964set}'s asymmetric filters (used
in the X-13ARIMA seasonal adjustment algorithm).

The initial series is modelled here as:
\begin{equation}\protect\hypertarget{eq-lpgeneralmodel}{}{
\boldsymbol y=\boldsymbol U \boldsymbol \gamma+\boldsymbol Z\boldsymbol \delta+\boldsymbol \varepsilon,\quad
\boldsymbol \varepsilon\sim\mathcal{N}(0,\boldsymbol D).
}\label{eq-lpgeneralmodel}\end{equation} where
\([\boldsymbol U,\boldsymbol Z]\) is of full rank and forms a subset of
the columns of \(X\). The objective is to find a filter
\(\boldsymbol v\) that minimises the mean square error of revision (to
the symmetric filter \(\boldsymbol \theta\)) under certain constraints.
These constraints are represented by the matrix
\(\boldsymbol U=\begin{pmatrix}\boldsymbol U_{p}\\\boldsymbol U_{f}\end{pmatrix}\):
\(\boldsymbol U_p'\boldsymbol v=\boldsymbol U'\boldsymbol \theta\) (with
\(\boldsymbol U_p\) the matrix \((h+q+1)\times (d+1)\) which contains
the observations of the matrix \(U\) known during estimation by the
asymmetric filter). The problem is equivalent to finding \(v\) which
minimises: \begin{equation}\protect\hypertarget{eq-lppasym}{}{
\varphi(\boldsymbol v)=
\underbrace{
  \underbrace{(\boldsymbol v-\boldsymbol \theta_{p})'\boldsymbol D_{p}(\boldsymbol v-\boldsymbol \theta_{p})+
  \boldsymbol \theta_{f}'\boldsymbol D_{f}\boldsymbol \theta_{f}}_\text{revision error variance}+
  \underbrace{[\boldsymbol \delta'(\boldsymbol Z_{p}'\boldsymbol v-\boldsymbol Z'\boldsymbol \theta)]^{2}}_{bias^2}
}_\text{mean square revision error}+
\underbrace{2\boldsymbol l'(\boldsymbol U_{p}'\boldsymbol v-\boldsymbol U'\boldsymbol \theta)}_{\text{constraints}}.
}\label{eq-lppasym}\end{equation} where \(\boldsymbol l\) is a vector of
Lagrange multipliers.

When \(\boldsymbol U=\boldsymbol X\), the constraint is equivalent to
preserving polynomials of degree \(d\): we find asymmetric direct
filters (DAF) when \(\boldsymbol D=\boldsymbol K^{-1}\).

When \(\boldsymbol U=\begin{pmatrix}1\\\vdots\\1\end{pmatrix}\),
\(\boldsymbol Z=\begin{pmatrix}-h\\\vdots\\+h\end{pmatrix}\),
\(\boldsymbol \delta=\delta_1\), \(\boldsymbol D=\sigma^2\boldsymbol I\)
and when the symmetric filter is the Henderson filter, we find the
asymmetric Musgrave filters (using the Henderson kernel). This filter
assumes that, for real-time estimation, the data is generated by a
linear process and that the asymmetric filters preserve the constants
(\(\sum v_i=\sum \theta_i=1\)). These asymmetric filters depend on the
ratio \(\lvert\delta_1/\sigma\rvert\), which, assuming that the trend is
linear and the bias constant, can be linked to the I-C ratio
\(R=\frac{\bar{I}}{\bar{C}}=\frac{\sum\lvert I_t-I_{t-1}\rvert}{\sum\lvert C_t-C_{t-1}\rvert}\):
\(\delta_1/\sigma=2/(R\sqrt{\pi})\). This ratio is used in particular in
X-13ARIMA to determine the length of the Henderson filter. For monthly
data:

\begin{itemize}
\item
  If the ratio is large (\(3.5< R\)) then a 23 term filter is used (to
  remove more noise).
\item
  If the ratio is small (\(R<1\)) a 9 term filter is used.
\item
  Otherwise (most of the cases) a 13-term filter is used.
\end{itemize}

When \(\boldsymbol U\) corresponds to the first \(d^*+1\) columns of
\(\boldsymbol X\), \(d^*<d\), the constraint is to reproduce polynomial
trends of degree \(d^*\). This introduces bias but reduces the variance.
Thus, \textcite{proietti2008} proposes three classes of asymmetric
filters:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \emph{Linear-Constant} (LC): \(y_t\) linear (\(d=1\)) and
  \(\boldsymbol v\) preserves constant signals (\(d^*=0\)). We obtain
  Musgrave filters when the Henderson kernel is used and the Henderson
  filter is used as the symmetric filter.
\item
  \emph{Quadratic-Linear} (QL): \(y_t\) quadratic (\(d=2\)) and
  \(\boldsymbol v\) preserves linear signals (\(d^*=1\)).
\item
  \emph{Cubic-Quadratic} (CQ): \(y_t\) cubic (\(d=3\)) and
  \(\boldsymbol v\) preserves quadratic signals (\(d^*=2\)).
\end{enumerate}

Appendix \ref{sec-an-cof} shows the coefficients, gain and phase shift
functions of the four asymmetric filters.

Table~\ref{tab-criteriaLp} compares the quality criteria of the
different methods using the Henderson filter and \(h=6\) (a symmetric
filter with 13 terms). For real-time filters (\(q=0\)), a more complex
asymmetrical filter (in terms of polynomial preservation) results in
lower timeliness and higher fidelity/smoothness: reduction in phase
shift comes at the expense of an increase in variance. This result
varies as \(q\) increases: for \(q=2\), the QL filter has greater
timeliness than the LC filter. This surprising result highlights the
fact that the phase shift is not controlled by \textcite{proietti2008}'s
approach.

In terms of revision (\(A_w+S_w+T_w+R_w\)), the LC and QL filters
consistently outperform the CQ and DAF filters.

These theoretical properties are consistent with the empirical results
observed in Section~\ref{sec-comparison}. Revisions are more important
for the CQ and DAF filters, which leads to greater variability in
estimations and a longer delay in detecting turning points. The globally
parameterised QL filter (with \(R\) fixed for all asymmetric filters)
may result in later detection of turning points than the LC filter.
Indeed, to identify a turning point at date \(t\), it is necessary to
have knowledge of at least two points after the given date to confirm
the reversal in trend.

\begin{table}[!h]
\caption{\label{tab-criteriaLp} Quality criteria for asymmetric filters ($q=0,1,2$) computed by local polynomials using the Henderson kernel with $h=6$ and $R=3,5$.}
\centering
\begin{tabular}[t]{cccccccccccc}
\toprule
Method & $b_c$ & $b_l$ & $b_q$ & $F_g$ & $S_g$ & $T_g \times 10^{-3}$ & $A_w$ & $S_w$ & $T_w$ & $R_w$ & $EQM_w$\\
\midrule
\addlinespace[0.3em]
\multicolumn{12}{l}{\textbf{$q=0$}}\\
\hspace{1em}LC & 0 & -0.41 & -2.16 & 0.39 & 1.27 & 30.34 & 0.10 & 0.49 & 0.41 & 0.55 & 1.54\\
\hspace{1em}QL & 0 & 0.00 & -0.47 & 0.71 & 5.15 & 0.05 & 0.07 & 1.89 & 0.00 & 0.11 & 2.07\\
\hspace{1em}CQ & 0 & 0.00 & 0.00 & 0.91 & 11.94 & 0.01 & 0.02 & 2.23 & 0.00 & 0.10 & 2.35\\
\hspace{1em}DAF & 0 & 0.00 & 0.00 & 0.94 & 14.20 & 0.00 & 0.01 & 2.18 & 0.00 & 0.10 & 2.29\\
\addlinespace[0.3em]
\multicolumn{12}{l}{\textbf{$q=1$}}\\
\hspace{1em}LC & 0 & -0.12 & -0.52 & 0.27 & 0.43 & 4.80 & 0.01 & 0.12 & 0.06 & 0.11 & 0.30\\
\hspace{1em}QL & 0 & 0.00 & -0.06 & 0.29 & 0.71 & 0.69 & 0.00 & 0.19 & 0.01 & 0.04 & 0.25\\
\hspace{1em}CQ & 0 & 0.00 & 0.00 & 0.37 & 0.57 & 0.16 & 0.02 & 0.58 & 0.00 & 0.06 & 0.66\\
\hspace{1em}DAF & 0 & 0.00 & 0.00 & 0.41 & 0.37 & 0.06 & 0.02 & 0.76 & 0.00 & 0.06 & 0.84\\
\addlinespace[0.3em]
\multicolumn{12}{l}{\textbf{$q=2$}}\\
\hspace{1em}LC & 0 & 0.00 & 1.08 & 0.20 & 0.08 & 0.35 & 0.01 & 0.01 & 0.00 & 0.01 & 0.04\\
\hspace{1em}QL & 0 & 0.00 & 0.03 & 0.22 & 0.05 & 2.08 & 0.00 & 0.01 & 0.02 & 0.07 & 0.10\\
\hspace{1em}CQ & 0 & 0.00 & 0.00 & 0.37 & 0.66 & 0.13 & 0.02 & 0.56 & 0.00 & 0.06 & 0.64\\
\hspace{1em}DAF & 0 & 0.00 & 0.00 & 0.40 & 0.77 & 0.02 & 0.02 & 0.68 & 0.00 & 0.05 & 0.74\\
\bottomrule
\end{tabular}

\textit{Note: With $EQM_w=A_w + S_w + T_w + R_w$.}
\end{table}

\hypertarget{sec-lptimeliness}{%
\subsection{Extension with the timeliness
criterion}\label{sec-lptimeliness}}

One drawback of the previous method is the lack of control over the
phase shift. However, it is possible to improve the modelling by
incorporating in Equation~\ref{eq-lppasym} the timeliness criterion
defined by \textcite{ch15HBSA}. This was proposed by Jean Palate, then
coded in Java and integrated into \texttt{rjd3filters}.

Using the same notation as in Section~\ref{sec-lppasymf},
\(\boldsymbol\theta\) represents the symmetric filter and
\(\boldsymbol v\) the asymmetric filter. With
\(\boldsymbol\theta=\begin{pmatrix}\boldsymbol\theta_p\\\boldsymbol\theta_f\end{pmatrix}\)
and \(\boldsymbol\theta_p\) of the same length of \(\boldsymbol v\), and
\(\boldsymbol g=\boldsymbol v-\boldsymbol \theta_p\), the timeliness
criterion can be expressed as: \[
T_g(\boldsymbol v)=\boldsymbol v'\boldsymbol T\boldsymbol v=\boldsymbol g'\boldsymbol T\boldsymbol g+2\boldsymbol \theta_p'\boldsymbol T\boldsymbol g+\boldsymbol \theta_p'\boldsymbol T\boldsymbol \theta_p
\quad\text{ with }\boldsymbol T\text{ a symmetric matrix}.
\] Furthermore, the objective function \(\varphi\) of
Equation~\ref{eq-lppasym} can be rewritten: \begin{align*}
\varphi(\boldsymbol v)&=(\boldsymbol v-\boldsymbol \theta_p)'\boldsymbol D_{p}(\boldsymbol v-\boldsymbol \theta_p)+
  \boldsymbol \theta_f'\boldsymbol D_{f}\boldsymbol \theta_f+
  [\boldsymbol \delta'(\boldsymbol Z_{p}'\boldsymbol v-\boldsymbol Z'\boldsymbol \theta)]^{2}+
2\boldsymbol l'(\boldsymbol U_{p}'\boldsymbol v-\boldsymbol U'\boldsymbol \theta)\\
&=\boldsymbol g'\boldsymbol Q\boldsymbol g-2\boldsymbol P\boldsymbol g+2\boldsymbol l'(\boldsymbol U_{p}'\boldsymbol v-\boldsymbol U'\boldsymbol \theta)+\boldsymbol c\quad\text{ with }\boldsymbol v
\begin{cases}
\boldsymbol Q=\boldsymbol D_p+\boldsymbol Z_p\boldsymbol \delta\boldsymbol \delta'\boldsymbol Z'_p \\
\boldsymbol P=\boldsymbol \theta_f\boldsymbol Z_f\boldsymbol \delta\boldsymbol \delta'\boldsymbol Z_p'\\
\boldsymbol c\text{ a constant independent of }\boldsymbol v
\end{cases}.
\end{align*} Adding the timeliness criterion: \[
\widetilde\varphi(\boldsymbol v)=\boldsymbol g'\widetilde {\boldsymbol Q}\boldsymbol g-
2\widetilde {\boldsymbol P}\boldsymbol g+2\boldsymbol l'(\boldsymbol U_{p}'\boldsymbol v-\boldsymbol U'\boldsymbol \theta)+
\widetilde {\boldsymbol c}\quad\text{with }
\begin{cases}
\widetilde {\boldsymbol Q}=\boldsymbol D_p+\boldsymbol Z_p\boldsymbol \delta\boldsymbol \delta'\boldsymbol Z'_p + \alpha_T\boldsymbol T\\
\widetilde {\boldsymbol P}=\boldsymbol \theta_f\boldsymbol Z_f\boldsymbol \delta\boldsymbol \delta'\boldsymbol Z_p'-\alpha_T\boldsymbol \theta_p\boldsymbol T\\
\widetilde {\boldsymbol c}\text{ a constant independent of }\boldsymbol v
\end{cases},
\] where \(\alpha_T\) is the weight associated to the timeliness
criterion. With \(\alpha_T=0\) we find \(\varphi(\boldsymbol v)\). This
extension therefore makes it possible to find all the symmetric and
asymmetric filters presented in the previous section but also
generalizes the approach of \textcite{GrayThomson1996}. This extension
is implemented in the \faIcon{r-project} function
\texttt{rjd3filters::lp\_filter()}.

One drawback is that \(T_g(\boldsymbol v)\) is not normalized: the
weight \(\alpha_T\) has no economic sense. It is then difficult to
calibrate the value of \(\alpha_T\) since the value of
\(\lvert\delta/\sigma\rvert\) must also be defined. That's why in this
paper we will not compare this extension to the other methods and we
will focus on the local calibration of \(\lvert\delta/\sigma\rvert\).
However, in future studies, we could imagine a two step calibration: fix
the value of \(\lvert\delta/\sigma\rvert\) to find \(\alpha_T\) by
cross-validation and then use this weight with a local parameterisation
of the ratio \(\lvert\delta/\sigma\rvert\) (as in
Section~\ref{sec-localic}).

\hypertarget{sec-localic}{%
\subsection{Local parameterisation of asymmetric
filters}\label{sec-localic}}

Asymmetric filters are usually parameterised globally:
\(\lvert\delta/\sigma\rvert\) estimated on all the data using the
IC-ratio or a cross-validation criterion. However, we might prefer a
local parameterisation: ratio \(\lvert\delta/\sigma\rvert\) which varies
as a function of time. Indeed, although the overall parameterisation is
generally valid, assuming a constant value of the
\(\lvert\delta/\sigma\rvert\) ratio for all asymmetrical filters does
not appear relevant for real-time estimation, especially during periods
of economic downturn. For example, with the LC method, global
parameterisation means assuming that the slope of the trend is constant,
whereas during economic downturns it tends towards 0 up to the turning
point.

This is what is proposed in this article, with a local parameterisation
of the asymmetric filters by estimating \(\delta\) and \(\sigma^2\)
separately. Although this does not give an unbiased estimator of the
ratio \(\lvert\delta/\sigma\rvert\), it does allow the main evolutions
to be captured, such as the decay towards 0 before a turning point and
the growth after the turning point for the LC method:

\begin{itemize}
\tightlist
\item
  The variance \(\sigma^2\) can be estimated using the observed data set
  and the symmetrical filter \((w_{-p},\dots,w_p)\): \[
  \hat\sigma^2=\frac{1}{n-2h}\sum_{t=h+1}^{n-h}\frac{(y_t-\hat \mu_t)^2}{1-2w_0^2+\sum w_i^2}.
  \]
\item
  The parameter \(\delta\) can be estimated by moving average from
  Equation~\ref{eq-mmsym}. For example, for the LC method we can use the
  moving average
  \(\boldsymbol \theta_2=\boldsymbol K\boldsymbol X(\boldsymbol X'\boldsymbol K\boldsymbol X)^{-1}\boldsymbol e_{2}\)
  to obtain a local estimate of the slope and for the QL method we can
  use
  \(\boldsymbol \theta_3=\boldsymbol K\boldsymbol X(\boldsymbol X'\boldsymbol K\boldsymbol X)^{-1}\boldsymbol e_{3}\)
  to obtain a local estimate of the concavity. The DAF method then
  simply allows us to calculate the associated asymmetric moving
  averages.\\
  Although a moving average of different length to that used for
  estimating the trend could be considered, this seems to degrade the
  results in terms of phase shift (using the same methodology as in
  Section~\ref{sec-comparison}). Furthermore, for the construction of
  moving averages, the trend can be modelled as being locally of degree
  2 or 3 (this has no impact on the final estimate of concavity). Here,
  we have chosen to model a trend of degree 2: this reduces the phase
  shift but slightly increases the revisions linked to the first
  estimate of the trend-cycle. Figure~\ref{fig-mmpenteconcac} shows the
  moving averages used.
\end{itemize}

\begin{figure}[H]

\caption{\label{fig-mmpenteconcac}Moving averages used for real-time
estimation of slope and concavity.}

{\centering \includegraphics[width=0.9\textwidth,height=\textheight]{img/filters_used/mm_penteconcavite.pdf}

}

\end{figure}

In the empirical applications of Section~\ref{sec-comparison}, the final
local parameterisation is obtained by estimating \(\delta\) using all
the data (i.e., using the symmetric filters shown in
Figure~\ref{fig-mmpenteconcac}), while maintaining a real-time estimate
of \(\sigma^2\). Figure~\ref{fig-mmpenteconcac-ex} shows an example of
the comparison of estimates of the ratio \(|\delta_1/\sigma|\) with a
global estimator (IC ratio) and two the local parameterisation:

\begin{itemize}
\item
  Figure~\ref{fig-mmpenteconcac-ex-1}: using a symmetric filter to
  estimate \(\delta_1\) (final local parameterisation).
\item
  Figure~\ref{fig-mmpenteconcac-ex-2}: using the asymmetric filter which
  need two points in the futur to estimate \(\delta_1\). This is closed
  to the real-time estimates since two points in the future are needed
  to detect a turning point.
\end{itemize}

The turning points are clearly detected by the local estimators, but
there is much more variance in the estimate of \(|\delta_1/\sigma|\) in
the real-time estimate. This will lead to another source of revisions in
the intermediate estimates.

\begin{figure}

\caption{\label{fig-mmpenteconcac-ex}Comparison of the estimates of
\(|\delta_1/\sigma|\) with a local estimator and global estimator (with
IC ratio) on a simulated series, the vertical lines being the simulated
turning points.}\begin{minipage}[t]{\linewidth}
\subcaption{\label{fig-mmpenteconcac-ex-1}Symmetric moving average used
for \(\hat \delta_1\).}

{\centering

\raisebox{-\height}{

\includegraphics{img/filters_used/mm_penteconcavite_ex.pdf}

}

}

\end{minipage}%
\newline
\begin{minipage}[t]{\linewidth}
\subcaption{\label{fig-mmpenteconcac-ex-2}Asymmetric moving average
using two futur points (\(q=2\)) used for \(\hat \delta_1\).}

{\centering

\raisebox{-\height}{

\includegraphics{img/filters_used/mm_penteconcavite_ex2.pdf}

}

}

\end{minipage}%

\end{figure}

\hypertarget{sec-comparison}{%
\section{Comparison of different methods}\label{sec-comparison}}

The different methods are compared on simulated and real data.

For all series, a symmetrical 13-term filter is used. These methods are
also compared with estimates obtained by extending the series with an
ARIMA model, then applying a 13-term symmetrical Henderson filter. The
ARIMA model is determined automatically, using no seasonal lag (the
series being seasonally adjusted) and no external variables (such as
external regressors for correction of atypical points).

The performance of the different methods is judged on criteria relating
to revisions (between two consecutive estimates and in relation to the
final estimate) and the number of periods required to detect turning
points.

\hypertarget{simulated-series}{%
\subsection{Simulated series}\label{simulated-series}}

\hypertarget{methodology}{%
\subsubsection{Methodology}\label{methodology}}

Following a methodology close to that of \textcite{DarneDagum2009}, nine
monthly series are simulated between January 1960 and December 2020 with
different levels of variability. Each simulated series
\(y_t= C_t+ T_t + I_t\) can be written as the sum of three components:

\begin{itemize}
\item
  the cycle
  \(C_t = \rho [\cos (2 \pi t / \lambda) +\sin (2 \pi t / \lambda)]\),
  \(\lambda\) is fixed at 72 (6-year cycles, so there are 19 detectable
  turning points);
\item
  the trend \(T_t = T_{t-1} + \nu_t\) with
  \(\nu_t \sim \mathcal{N}(0, \sigma_\nu^2)\), \(\sigma_\nu\) being
  fixed at \(0.08\);
\item
  and the irregular \(I_t = e_t\) with
  \(e_t \sim \mathcal{N}(0, \sigma_e^2)\).
\end{itemize}

For the different simulations, we vary the parameters \(\rho\) and
\(\sigma_e^2\) in order to obtain series with different signal-to-noise
ratios (see Figure~\ref{fig-graphs-data-simul}):

\begin{itemize}
\item
  High signal-to-noise ratio (i.e.~low I-C ratio and low variability):
  \(\sigma_e^2=0.2\) and \(\rho = 3.0, 3.5\) or \(4.0\) (I-C ratio
  between 0.9 and 0.7);
\item
  Medium signal-to-noise ratio (i.e.~medium I-C ratio and medium
  variability): \(\sigma_e^2=0.3\) and \(\rho = 1.5,\, 2.0\) or \(3.0\)
  (I-C ratio between 2.3 and 1.4);
\item
  Low signal-to-noise ratio (i.e.~high I-C ratio and high variability):
  \(\sigma_e^2=0.4\) and \(\rho = 0.5,\, 0.7\) or \(1.0\) (I-C ratio
  between 8.9 and 5.2).
\end{itemize}

\begin{figure}[H]

\caption{\label{fig-graphs-data-simul}Simulated series with low
(\(\sigma_e^2=0.2\) and \(\rho = 3,5\)), medium (\(\sigma_e^2=0.3\) and
\(\rho = 2.0\)) and high variability (\(\sigma_e^2=0.4\) and
\(\rho = 1.0\)).}

{\centering \includegraphics[width=1\textwidth,height=\textheight]{img/simulations/simul_data.pdf}

}

\end{figure}

For each series and each date, the trend-cycle is estimated using the
different methods presented in this report. Three quality criteria are
also calculated:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Computation of the phase shift in the detection of turning points. The
  definition of \textcite{Zellner1991} is used to determine the turning
  points:

  \begin{itemize}
  \tightlist
  \item
    An upturn occurs when the economy moves from a phase of recession to
    a phase of expansion. This is the case at date \(t\) when
    \(y_{t-3}\geq y_{t-2}\geq y_{t-1}<y_t\leq y_{t+1}\).\\
  \item
    A downturn occurs when the economy moves from a phase of expansion
    to a phase of recession. This is the case at date \(t\) when
    \(y_{t-3}\leq y_{t-2}\leq y_{t-1}>y_t\geq y_{t+1}\).\\
    It thus takes at least 2 months to detect a turning point.\\
    The phase shift is often defined as the number of months required to
    detect the right turning point. Here we use a slightly modified
    criterion: the phase shift is defined as the number of months needed
    to detect the right turning point without any future revision.
    Indeed, it can happen that the right turning point is detected by
    asymmetric filters but is not detected by the final estimate using a
    symmetric filter (this is the case for 41 reversal points out of the
    9 series with asymmetric Musgrave filters) or that there are
    revisions in successive estimates (this is the case for 7 turning
    points out of the 9 series with asymmetric Musgrave filters).
    Finally, relatively few turning points are detected at the right
    date with the final estimate. With the 13-term Henderson filter, 18
    are correctly detected in series with low variability (out of 57
    possible), 11 in series with medium variability and 12 in series
    with high variability.
  \end{itemize}
\item
  Computation of two revision criteria: the average of the relative
  deviations between the \(q\)\textsuperscript{th} estimate and the last
  estimate \(MAE_{fe}(q)\) and the average of the relative deviations
  between the \(q\)\textsuperscript{th} estimate and the
  \(q+1\)\textsuperscript{th} estimate \(MAE_{qe}(q)\). \[
  MAE_{fe}(q)=\mathbb E\left[
  \left|\frac{
  y_{t|t+q} -  y_{t|last}
  }{
   y_{t|last}
  }\right|
  \right]
  \quad\text{and}\quad
  MAE_{qe}(q)=\mathbb E\left[
  \left|\frac{
  y_{t|t+q} - y_{t|t+q+1}
  }{
  y_{t|t+q+1}
  }\right|
  \right].
  \]
\end{enumerate}

\hypertarget{comparison}{%
\subsection{Comparison}\label{comparison}}

Excluding for the moment the local parameterisations of the polynomial
filters, the linear-constant polynomial filter (LC) seems to give the
best results in terms of delay in the detection of turning points
(Figure~\ref{fig-graphstpsimul}). Performance is relatively close to
that obtained by extending the series using an ARIMA model. However,
when variability is low, the LC filter seems to give poorer results and
the quadratic-linear polynomial filter (QL) seems to give the best
results.

For series with moderate variability, local parameterisation of the LC
and QL filters reduces the phase shift. For series with high
variability, the phase shift is only reduced by using the final
\(\hat\delta\) parameters: real-time estimates seem to add more
variance. For series with low variability, performance seems to be
slightly improved only with the LC filter.

\begin{figure}[H]

\caption{\label{fig-graphstpsimul}Distribution of phase shifts on
simulated series.}

{\centering \includegraphics[width=1\textwidth,height=\textheight]{img/simulations/phase_shift_simul.pdf}

}

\end{figure}

In terms of revisions, the variability of the series has little effect
on the respective performances of the different methods, but it does
affect the orders of magnitude, which is why the results are presented
only for series with medium variability (Table~\ref{tab-simulrev}). In
general, LC filters always minimise revisions (with relatively small
effect from the local parameterisation of the filters) and revisions are
greater with cubic-quadratic (CQ) and direct (DAF) polynomial filters.

For the QL filter, there is a large revision between the second and
third estimates: this may be due to the fact that for the second
estimate (when one point in the future is known), the QL filter assigns
a greater weight to the estimate in \(t+1\) than to the estimate in
\(t\), which creates a discontinuity. This revision is considerably
reduced by parameterising the filter locally. For polynomial filters
other than the LC filter, the large revisions at the first estimate were
to be expected given the coefficient curve: a very large weight is
associated with the current observation and there is a strong
discontinuity between the moving average used for real-time estimation
(when no point in the future is known) and the other moving averages.

Extending the series using an ARIMA model gives revisions with the
latest estimates of the same order of magnitude as the LC filter, but
slightly larger revisions between consecutive estimates, particularly
between the fourth and fifth estimates (as might be expected, as
highlighted in Section~\ref{sec-mmetprev}).

\begin{table}[!h]
\caption{\label{tab-simulrev} Average of the relative deviations of the revisions for the different
filters on simulated series with medium variability.}
\centering
\begin{tabular}{ccccccc}
\toprule
Method & $q=0$ & $q=1$ & $q=2$ & $q=3$ & $q=4$ & $q=5$\\
\midrule
\addlinespace[0.3em]
\multicolumn{7}{l}{\textbf{$MAE_{fe}(q) = \mathbb E\left[\left|(y_{t|t+q} -  y_{t|last})/y_{t|last}\right|\right]$}}\\
\hspace{1em}LC & 0.21 & 0.10 & 0.03 & 0.03 & 0.03 & 0.01\\
\hspace{1em}LC local param. (final estimates) & 0.19 & 0.09 & 0.03 & 0.03 & 0.03 & 0.01\\
\hspace{1em}LC local param. & 0.29 & 0.10 & 0.03 & 0.03 & 0.03 & 0.01\\
\hspace{1em}QL & 0.33 & 0.10 & 0.04 & 0.04 & 0.03 & 0.01\\
\hspace{1em}QL local param. (final estimates) & 0.21 & 0.10 & 0.03 & 0.03 & 0.03 & 0.01\\
\hspace{1em}QL local param. & 0.30 & 0.10 & 0.04 & 0.03 & 0.03 & 0.01\\
\hspace{1em}CQ & 0.45 & 0.13 & 0.13 & 0.09 & 0.06 & 0.02\\
\hspace{1em}DAF & 0.47 & 0.15 & 0.15 & 0.09 & 0.06 & 0.02\\
\hspace{1em}ARIMA & 0.22 & 0.10 & 0.03 & 0.03 & 0.03 & 0.01\\
\addlinespace[0.3em]
\multicolumn{7}{l}{\textbf{$MAE_{ce}(q)=\mathbb E\left[
\left|(y_{t|t+q} - y_{t|t+q+1})/y_{t|t+q+1}\right|
\right]$}}\\
\hspace{1em}LC & 0.19 & 0.10 & 0.02 & 0.01 & 0.07 & 0.01\\
\hspace{1em}LC local param. (final estimates) & 0.20 & 0.10 & 0.03 & 0.01 & 0.05 & 0.01\\
\hspace{1em}LC local param. & 0.24 & 0.11 & 0.03 & 0.01 & 0.05 & 0.01\\
\hspace{1em}QL & 0.29 & 3.46 & 0.00 & 0.03 & 0.04 & 0.01\\
\hspace{1em}QL local param. (final estimates) & 0.31 & 0.11 & 0.02 & 0.01 & 0.04 & 0.01\\
\hspace{1em}QL local param. & 0.24 & 0.16 & 0.00 & 0.03 & 0.04 & 0.01\\
\hspace{1em}CQ & 0.43 & 0.02 & 0.10 & 0.07 & 0.05 & 0.02\\
\hspace{1em}DAF & 0.66 & 0.24 & 0.11 & 0.14 & 0.06 & 0.02\\
\hspace{1em}ARIMA & 0.21 & 0.13 & 0.02 & 0.02 & 0.25 & 0.01\\
\bottomrule
\end{tabular}
\end{table}

\hypertarget{real-series}{%
\subsection{Real series}\label{real-series}}

The differences between the methods are also illustrated using an
example from the FRED-MD database (\textcite{fredmd}) containing
economic series on the United States. The series studied correspond to
the database published in November 2022. The series studied is the level
of employment in the United States (series \texttt{CE160V}, used in
logarithm) around the February 2001 turning point, consistent with the
monthly dating of the turning points. This point and this series were
chosen because the economic turnaround is particularly visible in the
raw series (Figure~\ref{fig-ce16ov-previmp-lp}), because this series is
used for dating economic cycles and because the FRED-MD database
facilitates the reproducibility of the results (thanks to the
availability of series published on past dates). Studying the series up
to January 2020, this series has a medium variability: a symmetrical
filter of 13 terms is therefore appropriate. Appendix
\autocite*{ann-ex-r} shows how the code to reproduce the different
plots.

Figure~\ref{fig-ce16ovlp} shows the successive estimates of the
trend-cycle using the different methods studied.

For this series, the phase shift is 6 months for the LC and CQ methods
and the extension of the series by ARIMA. Except for the LC method
(where the turning point is first detected in January 2001), the high
phase shift is due to revisions in the turning point detected in the
intermediate estimates (even if the correct turning point is detected
after 2 months, it is no longer detected after 3 months). It is of two
months for the other methods (QL, DAF).

In this example, local parameterisation does not reduce the phase shift,
but it does reduce revisions. The QL and DAF polynomials lead to greater
variability in the intermediate estimates, especially in February 2001.

\begin{figure}[H]

\caption{\label{fig-ce16ovlp}Successive estimates of the trend-cycle in
US employment (in logarithms).}

{\centering \includegraphics[width=0.95\textwidth,height=\textheight]{img/nber/ce16ov_fev2001_lp.pdf}

}

\end{figure}

The quality of the intermediate estimates can also be analysed using the
implicit forecasts of the different methods
(Figure~\ref{fig-ce16ov-previmp-lp}). As a reminder, these are forecasts
of the raw series which, by applying the symmetrical Henderson filter to
the extended series, give the same estimates as the asymmetrical moving
averages. The forecasts of the ARIMA model are naive and do not take the
turning point into account, unlike the other methods. Finally, local
parameterisation of the QL filter produces much more consistent
forecasts.

\begin{figure}[H]

\caption{\label{fig-ce16ov-previmp-lp}Implicit forecasts linked to
successive estimates of the trend-cycle in US employment (in logarithms)
using local polynomial methods.}

{\centering \includegraphics[width=0.95\textwidth,height=\textheight]{img/nber/ce16ov_fev2001_prev_imp_lp.pdf}

}

\end{figure}

\hypertarget{conclusion}{%
\section*{Conclusion}\label{conclusion}}
\addcontentsline{toc}{section}{Conclusion}

For business cycle analysis, most statisticians use trend-cycle
extraction methods, either directly or indirectly. They are used, for
example, to reduce the noise of an indicator in order to improve its
analysis, and models, such as forecasting models, usually use seasonally
adjusted series based on these methods.

This paper presents the \faIcon{r-project} package \texttt{rjd3filters},
which implements a several methods to build moving average for real-time
trend-cycle estimation. It also offers various functions for the
theoretical analysis of moving averages (gain functions, phase shift,
quality criteria, etc.) and judging the quality of the latest estimates
(for example with implicit forecasts). It also makes it easy to combine
moving averages and also to integrate custom moving averages into the
X-11 seasonal adjustment algorithm (with the
\texttt{rjd3filters::x11plus\_trend()} function). \texttt{rjd3filters}
thus facilitates research into the use of moving averages as part of
real-time cycle trend estimation. This is the case, for example, with
the local parametrisation of Musgrave moving averages and other
asymmetric moving averages linked to local polynomial regression,
presented in this article.

By comparing the different methods, we can learn a few lessons about the
construction of these moving averages.

During economic downturns, asymmetric filters used as an alternative to
extending the series using the ARIMA model can reduce revisions to
intermediate estimates of the trend-cycle and enable turning points to
be detected more quickly.

At the end of the period, modelling polynomial trends of degree greater
than three (cubic-quadratic, CQ, and direct, DAF) seems to introduce
variance into the estimates (and therefore more revisions) without
allowing faster detection of turning points. At the end of the period,
for estimating the trend-cycle, we can therefore restrict ourselves to
methods modelling polynomial trends of degree two or less
(linear-constant, LC, and quadratic-linear, QL). In addition,
parameterising polynomial filters locally as suggested in this paper
enables turning points to be detected more quickly (especially for the
QL filter). Even when the phase shift is not reduced, local
parameterisation is recommended because it reduces revisions and
produces intermediate estimates that are more consistent with expected
future trends. However, with these methods, the length of the filter
used must be adapted to the variability of the series: if the filter
used is too short (i.e.~if the variability of the series is ``low''),
retaining polynomial trends of degree one or less (LC method) produces
poorer results in terms of detecting turning points.

This study could be extended in many ways.

One possible extension would be to look at the impact of filter length
on the detection of turning points. Asymmetric filters are calibrated
using indicators calculated for the estimation of symmetric filters (for
example, to automatically determine their length), whereas a local
estimate might be preferable. Furthermore, we have only focused on
monthly series with a 13-term symmetric filter, but the results may be
different if the symmetric filter studied is longer or shorter and if we
study series with other frequencies (quarterly or daily, for example).

Another possibility would be to study the impact of atypical points:
moving averages, like any linear operator, are very sensitive to the
presence of atypical points. To limit their impact, in X-13ARIMA a
strong correction for atypical points is performed on the irregular
component before applying the filters to extract the trend-cycle. This
leads us to study the impact of these points on the estimation of the
trend-cycle and turning points, and also to explore new types of
asymmetric filters based on robust methods (such as robust local
regressions or moving medians).

\newpage

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\printbibliography[heading=none]

\appendix
\newpage

\hypertarget{sec-an-cof}{%
\section{Coefficients, gain and phase shift
functions}\label{sec-an-cof}}

\begin{figure}[H]

\caption{\label{fig-graphs-coef-lc}Coefficients, gain and phase shift
functions for the Linear-Constant (LC) filter with \(I/C=3.5\).}

{\centering \includegraphics[width=1\textwidth,height=\textheight]{img/filters_used/lc.pdf}

}

\end{figure}

\begin{figure}[H]

\caption{\label{fig-graphs-coef-ql}Coefficients, gain and phase shift
functions for the Quadratic-Linear (QL) filter with \(I/C=3.5\).}

{\centering \includegraphics[width=1\textwidth,height=\textheight]{img/filters_used/ql.pdf}

}

\end{figure}

\begin{figure}[H]

\caption{\label{fig-graphs-coef-cq}Coefficients, gain and phase shift
functions for the Cubic-Quadratic (CQ) filter with \(I/C=3.5\).}

{\centering \includegraphics[width=1\textwidth,height=\textheight]{img/filters_used/cq.pdf}

}

\end{figure}

\begin{figure}[H]

\caption{\label{fig-graphs-coef-daf}Coefficients, gain and phase shift
functions for the direct asymmetric filter (DAF).}

{\centering \includegraphics[width=1\textwidth,height=\textheight]{img/filters_used/daf.pdf}

}

\end{figure}

\hypertarget{ann-ex-r}{%
\section{R code example}\label{ann-ex-r}}

In this appendix we present the \faIcon{r-project} code that can be used
to estimate trend-cycle components for the US Employment with local
polynomial filters.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# \# To install rjd3filters}
\CommentTok{\# remotes::install\_github("rjdemetra/rjd3toolkit")}
\CommentTok{\# remotes::install\_github("rjdemetra/rjd3x11plus")}
\CommentTok{\# remotes::install\_github("rjdemetra/rjd3filters")}
\FunctionTok{library}\NormalTok{(rjd3filters)}
\FunctionTok{library}\NormalTok{(ggplot2)}
\FunctionTok{library}\NormalTok{(patchwork)}
\FunctionTok{library}\NormalTok{(zoo)}
\FunctionTok{library}\NormalTok{(forecast)}
\NormalTok{file }\OtherTok{\textless{}{-}}\StringTok{"https://files.stlouisfed.org/files/htdocs/fred{-}md/monthly/2022{-}11.csv"}
\NormalTok{data }\OtherTok{\textless{}{-}} \FunctionTok{read.csv}\NormalTok{(file)}
\NormalTok{data }\OtherTok{\textless{}{-}}\NormalTok{ data[}\SpecialCharTok{{-}}\DecValTok{1}\NormalTok{,] }\CommentTok{\# First line removed: it contains the transformation codes}
\NormalTok{y }\OtherTok{\textless{}{-}} \FunctionTok{ts}\NormalTok{(}\FunctionTok{log}\NormalTok{(data[, }\StringTok{"CE16OV"}\NormalTok{]), }\CommentTok{\# We study US employment}
         \AttributeTok{start =} \DecValTok{1959}\NormalTok{, }\AttributeTok{frequency =} \DecValTok{12}\NormalTok{)}
\NormalTok{y }\OtherTok{\textless{}{-}} \FunctionTok{window}\NormalTok{(y, }\AttributeTok{end =} \FunctionTok{c}\NormalTok{(}\DecValTok{2001}\NormalTok{, }\DecValTok{9}\NormalTok{))}
\NormalTok{last\_dates }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\FunctionTok{tail}\NormalTok{(}\FunctionTok{time}\NormalTok{(y), }\DecValTok{8}\NormalTok{))}
\FunctionTok{names}\NormalTok{(last\_dates) }\OtherTok{\textless{}{-}} \FunctionTok{as.character}\NormalTok{(zoo}\SpecialCharTok{::}\FunctionTok{as.yearmon}\NormalTok{(last\_dates))}
\NormalTok{last\_est }\OtherTok{\textless{}{-}} \FunctionTok{lapply}\NormalTok{(last\_dates, }\ControlFlowTok{function}\NormalTok{(x) }\FunctionTok{window}\NormalTok{(y, }\AttributeTok{end =}\NormalTok{ x))}

\CommentTok{\# MA of length 13 is appropriated:}
\FunctionTok{select\_trend\_filter}\NormalTok{(y)}
\CommentTok{\# Final estimates:}
\NormalTok{tc\_f }\OtherTok{\textless{}{-}} \FunctionTok{henderson}\NormalTok{(y, }\AttributeTok{length =} \DecValTok{13}\NormalTok{, }\AttributeTok{musgrave =} \ConstantTok{FALSE}\NormalTok{)}
\FunctionTok{plot}\NormalTok{(y)}
\FunctionTok{lines}\NormalTok{(tc\_f, }\AttributeTok{col =} \StringTok{"red"}\NormalTok{)}
\NormalTok{forecast}\SpecialCharTok{::}\FunctionTok{autoplot}\NormalTok{(}\FunctionTok{ts.union}\NormalTok{(y, tc\_f))}

\CommentTok{\# IC{-}ratio computation}
\NormalTok{icr }\OtherTok{\textless{}{-}} \FunctionTok{sapply}\NormalTok{(last\_est, }\ControlFlowTok{function}\NormalTok{(x) \{}
  \FunctionTok{ic\_ratio}\NormalTok{(x, }\FunctionTok{henderson}\NormalTok{(x, }\AttributeTok{length =} \DecValTok{13}\NormalTok{, }\AttributeTok{musgrave =} \ConstantTok{FALSE}\NormalTok{))}
\NormalTok{\})}
\NormalTok{lp\_est }\OtherTok{\textless{}{-}} \FunctionTok{lapply}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\StringTok{"LC"}\NormalTok{, }\StringTok{"QL"}\NormalTok{, }\StringTok{"CQ"}\NormalTok{, }\StringTok{"DAF"}\NormalTok{), }\ControlFlowTok{function}\NormalTok{(method) \{}
\NormalTok{  res }\OtherTok{\textless{}{-}} \FunctionTok{lapply}\NormalTok{(}\FunctionTok{seq\_along}\NormalTok{(icr), }\ControlFlowTok{function}\NormalTok{(i) \{}
\NormalTok{    lp\_coef }\OtherTok{\textless{}{-}} \FunctionTok{lp\_filter}\NormalTok{(}\AttributeTok{horizon =} \DecValTok{6}\NormalTok{,}
                         \AttributeTok{kernel =} \StringTok{"Henderson"}\NormalTok{,}
                         \AttributeTok{endpoints =}\NormalTok{ method,}
                         \AttributeTok{ic =}\NormalTok{ icr[i])}
\NormalTok{    rjd3filters}\SpecialCharTok{::}\FunctionTok{filter}\NormalTok{(last\_est[[i]], lp\_coef)}
\NormalTok{  \})}
  \FunctionTok{names}\NormalTok{(res) }\OtherTok{\textless{}{-}} \FunctionTok{names}\NormalTok{(last\_est)}
\NormalTok{  res}
\NormalTok{\})}
\NormalTok{lp\_if }\OtherTok{\textless{}{-}} \FunctionTok{lapply}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\StringTok{"LC"}\NormalTok{, }\StringTok{"QL"}\NormalTok{, }\StringTok{"CQ"}\NormalTok{, }\StringTok{"DAF"}\NormalTok{), }\ControlFlowTok{function}\NormalTok{(method) \{}
\NormalTok{  res }\OtherTok{\textless{}{-}} \FunctionTok{lapply}\NormalTok{(}\FunctionTok{seq\_along}\NormalTok{(icr), }\ControlFlowTok{function}\NormalTok{(i) \{}
\NormalTok{    lp\_coef }\OtherTok{\textless{}{-}} \FunctionTok{lp\_filter}\NormalTok{(}\AttributeTok{horizon =} \DecValTok{6}\NormalTok{,}
                         \AttributeTok{kernel =} \StringTok{"Henderson"}\NormalTok{,}
                         \AttributeTok{endpoints =}\NormalTok{ method,}
                         \AttributeTok{ic =}\NormalTok{ icr[i])}
    \FunctionTok{implicit\_forecast}\NormalTok{(last\_est[[i]], lp\_coef)}
\NormalTok{  \})}
  \FunctionTok{names}\NormalTok{(res) }\OtherTok{\textless{}{-}} \FunctionTok{names}\NormalTok{(last\_est)}
\NormalTok{  res}
\NormalTok{\})}
\FunctionTok{names}\NormalTok{(lp\_est) }\OtherTok{\textless{}{-}} \FunctionTok{names}\NormalTok{(lp\_if) }\OtherTok{\textless{}{-}}  \FunctionTok{c}\NormalTok{(}\StringTok{"LC"}\NormalTok{, }\StringTok{"QL"}\NormalTok{, }\StringTok{"CQ"}\NormalTok{, }\StringTok{"DAF"}\NormalTok{)}

\CommentTok{\# Local estimates of IC{-}ratios}
\CommentTok{\# We replicate the direct estimates to have}
\CommentTok{\# estimators of the slope and the concavity}
\NormalTok{gen\_MM }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(}\AttributeTok{p=}\DecValTok{6}\NormalTok{, }\AttributeTok{q=}\NormalTok{p, }\AttributeTok{d=}\DecValTok{2}\NormalTok{)\{}
\NormalTok{  X\_gen }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(}\AttributeTok{d =} \DecValTok{1}\NormalTok{, }\AttributeTok{p =} \DecValTok{6}\NormalTok{, }\AttributeTok{q =}\NormalTok{ p)\{}
    \FunctionTok{sapply}\NormalTok{(}\DecValTok{0}\SpecialCharTok{:}\NormalTok{d, }\ControlFlowTok{function}\NormalTok{(exp) }\FunctionTok{seq}\NormalTok{(}\SpecialCharTok{{-}}\NormalTok{p, q)}\SpecialCharTok{\^{}}\NormalTok{exp)}
\NormalTok{  \}}
\NormalTok{  k }\OtherTok{=}\NormalTok{ rjd3filters}\SpecialCharTok{::}\FunctionTok{get\_kernel}\NormalTok{(}\StringTok{"Henderson"}\NormalTok{, }\AttributeTok{h =}\NormalTok{ p)}
\NormalTok{  k }\OtherTok{=} \FunctionTok{c}\NormalTok{(}\FunctionTok{rev}\NormalTok{(k}\SpecialCharTok{$}\NormalTok{coef[}\SpecialCharTok{{-}}\DecValTok{1}\NormalTok{]), k}\SpecialCharTok{$}\NormalTok{coef[}\FunctionTok{seq}\NormalTok{(}\DecValTok{0}\NormalTok{,q)}\SpecialCharTok{+}\DecValTok{1}\NormalTok{])}
\NormalTok{  K }\OtherTok{=} \FunctionTok{diag}\NormalTok{(k)}
\NormalTok{  X }\OtherTok{=} \FunctionTok{X\_gen}\NormalTok{(}\AttributeTok{d=}\NormalTok{d, }\AttributeTok{p =}\NormalTok{ p, }\AttributeTok{q =}\NormalTok{ q)}
\NormalTok{  e1 }\OtherTok{=}\NormalTok{ e2 }\OtherTok{=}\NormalTok{ e3 }\OtherTok{=} \FunctionTok{matrix}\NormalTok{(}\DecValTok{0}\NormalTok{, }\AttributeTok{ncol =} \DecValTok{1}\NormalTok{, }\AttributeTok{nrow =}\NormalTok{ d}\SpecialCharTok{+}\DecValTok{1}\NormalTok{)}
\NormalTok{  e1[}\DecValTok{1}\NormalTok{] }\OtherTok{=} \DecValTok{1}
\NormalTok{  e2[}\DecValTok{2}\NormalTok{] }\OtherTok{=} \DecValTok{1}
\NormalTok{  e3[}\DecValTok{3}\NormalTok{] }\OtherTok{=} \DecValTok{1}
  \CommentTok{\# Estimator of the constant}
\NormalTok{  M1 }\OtherTok{=}\NormalTok{ K }\SpecialCharTok{\%*\%}\NormalTok{ X }\SpecialCharTok{\%*\%} \FunctionTok{solve}\NormalTok{(}\FunctionTok{t}\NormalTok{(X) }\SpecialCharTok{\%*\%}\NormalTok{ K }\SpecialCharTok{\%*\%}\NormalTok{ X, e1)}
  \CommentTok{\# Estimator of the slope}
\NormalTok{  M2 }\OtherTok{=}\NormalTok{ K }\SpecialCharTok{\%*\%}\NormalTok{ X }\SpecialCharTok{\%*\%} \FunctionTok{solve}\NormalTok{(}\FunctionTok{t}\NormalTok{(X) }\SpecialCharTok{\%*\%}\NormalTok{ K }\SpecialCharTok{\%*\%}\NormalTok{ X, e2)}
  \CommentTok{\# Estimor of the concavity}
\NormalTok{  M3 }\OtherTok{=}\NormalTok{ K }\SpecialCharTok{\%*\%}\NormalTok{ X }\SpecialCharTok{\%*\%} \FunctionTok{solve}\NormalTok{(}\FunctionTok{t}\NormalTok{(X) }\SpecialCharTok{\%*\%}\NormalTok{ K }\SpecialCharTok{\%*\%}\NormalTok{ X, e3)}
\NormalTok{  mm }\OtherTok{\textless{}{-}} \FunctionTok{list}\NormalTok{(}\AttributeTok{const =}\NormalTok{ M1, }\AttributeTok{slope =}\NormalTok{ M2, }\AttributeTok{concav =}\NormalTok{ M3)}
  \FunctionTok{lapply}\NormalTok{(mm, moving\_average, }\AttributeTok{lags =} \SpecialCharTok{{-}}\NormalTok{p)}
\NormalTok{\}}
\NormalTok{all\_mm }\OtherTok{\textless{}{-}} \FunctionTok{lapply}\NormalTok{(}\DecValTok{6}\SpecialCharTok{:}\DecValTok{0}\NormalTok{, gen\_MM, }\AttributeTok{p =} \DecValTok{6}\NormalTok{, }\AttributeTok{d =} \DecValTok{2}\NormalTok{)}
\NormalTok{est\_slope }\OtherTok{\textless{}{-}} \FunctionTok{finite\_filters}\NormalTok{(all\_mm[[}\DecValTok{1}\NormalTok{]]}\SpecialCharTok{$}\NormalTok{slope,}
                            \FunctionTok{lapply}\NormalTok{(all\_mm[}\SpecialCharTok{{-}}\DecValTok{1}\NormalTok{], }\StringTok{\textasciigrave{}}\AttributeTok{[[}\StringTok{\textasciigrave{}}\NormalTok{, }\StringTok{"slope"}\NormalTok{))}
\NormalTok{est\_concav }\OtherTok{\textless{}{-}} \FunctionTok{finite\_filters}\NormalTok{(all\_mm[[}\DecValTok{1}\NormalTok{]]}\SpecialCharTok{$}\NormalTok{concav,}
                             \FunctionTok{lapply}\NormalTok{(all\_mm[}\SpecialCharTok{{-}}\DecValTok{1}\NormalTok{], }\StringTok{\textasciigrave{}}\AttributeTok{[[}\StringTok{\textasciigrave{}}\NormalTok{, }\StringTok{"concav"}\NormalTok{))}

\NormalTok{henderson\_f }\OtherTok{\textless{}{-}} \FunctionTok{lp\_filter}\NormalTok{(}\AttributeTok{h=}\DecValTok{6}\NormalTok{)}\SpecialCharTok{@}\NormalTok{sfilter}
\NormalTok{lp\_filter2 }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(icr, }\AttributeTok{method =} \StringTok{"LC"}\NormalTok{, }\AttributeTok{h =} \DecValTok{6}\NormalTok{, }\AttributeTok{kernel =} \StringTok{"Henderson"}\NormalTok{)\{}
\NormalTok{  all\_coef }\OtherTok{=} \FunctionTok{lapply}\NormalTok{(icr, }\ControlFlowTok{function}\NormalTok{(ic)\{}
    \FunctionTok{lp\_filter}\NormalTok{(}\AttributeTok{horizon =}\NormalTok{ h,}
              \AttributeTok{kernel =}\NormalTok{ kernel,}
              \AttributeTok{endpoints =}\NormalTok{ method,}
              \AttributeTok{ic =}\NormalTok{ ic)}
\NormalTok{  \})}
\NormalTok{  rfilters }\OtherTok{=} \FunctionTok{lapply}\NormalTok{(}\DecValTok{1}\SpecialCharTok{:}\NormalTok{h, }\ControlFlowTok{function}\NormalTok{(i)\{}
\NormalTok{    q}\OtherTok{=}\NormalTok{h }\SpecialCharTok{{-}}\NormalTok{i}
\NormalTok{    all\_coef[[i]][,}\FunctionTok{sprintf}\NormalTok{(}\StringTok{"q=\%i"}\NormalTok{, q)]}
\NormalTok{  \})}
  \FunctionTok{finite\_filters}\NormalTok{(henderson\_f, }\AttributeTok{rfilters =}\NormalTok{ rfilters)}
\NormalTok{\}}
\NormalTok{loc\_lc\_est }\OtherTok{\textless{}{-}}
  \FunctionTok{lapply}\NormalTok{(last\_est, }\ControlFlowTok{function}\NormalTok{(x) \{}
\NormalTok{    est\_loc\_slope }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\FunctionTok{tail}\NormalTok{(est\_slope }\SpecialCharTok{*}\NormalTok{ x, }\DecValTok{6}\NormalTok{))}
\NormalTok{    sigma2 }\OtherTok{\textless{}{-}} \FunctionTok{var\_estimator}\NormalTok{(x, henderson\_f)}
\NormalTok{    icr }\OtherTok{=} \DecValTok{2}\SpecialCharTok{/}\NormalTok{(}\FunctionTok{sqrt}\NormalTok{(pi) }\SpecialCharTok{*}\NormalTok{ (est\_loc\_slope }\SpecialCharTok{/} \FunctionTok{sqrt}\NormalTok{(sigma2)))}
\NormalTok{    lp\_coef }\OtherTok{=} \FunctionTok{lp\_filter2}\NormalTok{(}\AttributeTok{ic =}\NormalTok{ icr,}
                         \AttributeTok{method =} \StringTok{"LC"}\NormalTok{, }\AttributeTok{h =} \DecValTok{6}\NormalTok{, }\AttributeTok{kernel =} \StringTok{"Henderson"}\NormalTok{)}
\NormalTok{    rjd3filters}\SpecialCharTok{::}\FunctionTok{filter}\NormalTok{(x, lp\_coef)}
\NormalTok{  \})}
\NormalTok{loc\_lc\_if }\OtherTok{\textless{}{-}}
  \FunctionTok{lapply}\NormalTok{(last\_est, }\ControlFlowTok{function}\NormalTok{(x) \{}
\NormalTok{    est\_loc\_slope }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\FunctionTok{tail}\NormalTok{(est\_slope }\SpecialCharTok{*}\NormalTok{ x, }\DecValTok{6}\NormalTok{))}
\NormalTok{    sigma2 }\OtherTok{\textless{}{-}} \FunctionTok{var\_estimator}\NormalTok{(x, henderson\_f)}
\NormalTok{    icr }\OtherTok{=} \DecValTok{2}\SpecialCharTok{/}\NormalTok{(}\FunctionTok{sqrt}\NormalTok{(pi) }\SpecialCharTok{*}\NormalTok{ (est\_loc\_slope }\SpecialCharTok{/} \FunctionTok{sqrt}\NormalTok{(sigma2)))}
\NormalTok{    lp\_coef }\OtherTok{=} \FunctionTok{lp\_filter2}\NormalTok{(}\AttributeTok{ic =}\NormalTok{ icr,}
                         \AttributeTok{method =} \StringTok{"LC"}\NormalTok{, }\AttributeTok{h =} \DecValTok{6}\NormalTok{, }\AttributeTok{kernel =} \StringTok{"Henderson"}\NormalTok{)}
    \FunctionTok{implicit\_forecast}\NormalTok{(x, lp\_coef)}
\NormalTok{  \})}
\NormalTok{loc\_ql\_est }\OtherTok{\textless{}{-}}
  \FunctionTok{lapply}\NormalTok{(last\_est, }\ControlFlowTok{function}\NormalTok{(x) \{}
\NormalTok{    est\_loc\_concav }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\FunctionTok{tail}\NormalTok{(est\_concav }\SpecialCharTok{*}\NormalTok{ x, }\DecValTok{6}\NormalTok{))}
\NormalTok{    sigma2 }\OtherTok{\textless{}{-}} \FunctionTok{var\_estimator}\NormalTok{(x, henderson\_f)}
\NormalTok{    icr }\OtherTok{=} \DecValTok{2}\SpecialCharTok{/}\NormalTok{(}\FunctionTok{sqrt}\NormalTok{(pi) }\SpecialCharTok{*}\NormalTok{ (est\_loc\_concav }\SpecialCharTok{/} \FunctionTok{sqrt}\NormalTok{(sigma2)))}
\NormalTok{    lp\_coef }\OtherTok{=} \FunctionTok{lp\_filter2}\NormalTok{(}\AttributeTok{ic =}\NormalTok{ icr,}
                         \AttributeTok{method =} \StringTok{"QL"}\NormalTok{, }\AttributeTok{h =} \DecValTok{6}\NormalTok{, }\AttributeTok{kernel =} \StringTok{"Henderson"}\NormalTok{)}
\NormalTok{    rjd3filters}\SpecialCharTok{::}\FunctionTok{filter}\NormalTok{(x, lp\_coef)}
\NormalTok{  \})}
\NormalTok{loc\_ql\_if }\OtherTok{\textless{}{-}}
  \FunctionTok{lapply}\NormalTok{(last\_est, }\ControlFlowTok{function}\NormalTok{(x) \{}
\NormalTok{    est\_loc\_concav }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\FunctionTok{tail}\NormalTok{(est\_concav }\SpecialCharTok{*}\NormalTok{ x, }\DecValTok{6}\NormalTok{))}
\NormalTok{    sigma2 }\OtherTok{\textless{}{-}} \FunctionTok{var\_estimator}\NormalTok{(x, henderson\_f)}
\NormalTok{    icr }\OtherTok{=} \DecValTok{2}\SpecialCharTok{/}\NormalTok{(}\FunctionTok{sqrt}\NormalTok{(pi) }\SpecialCharTok{*}\NormalTok{ (est\_loc\_concav }\SpecialCharTok{/} \FunctionTok{sqrt}\NormalTok{(sigma2)))}
\NormalTok{    lp\_coef }\OtherTok{=} \FunctionTok{lp\_filter2}\NormalTok{(}\AttributeTok{ic =}\NormalTok{ icr,}
                         \AttributeTok{method =} \StringTok{"QL"}\NormalTok{, }\AttributeTok{h =} \DecValTok{6}\NormalTok{, }\AttributeTok{kernel =} \StringTok{"Henderson"}\NormalTok{)}
    \FunctionTok{implicit\_forecast}\NormalTok{(x, lp\_coef)}
\NormalTok{  \})}

\DocumentationTok{\#\# Plots}
\CommentTok{\# Plots with all the estimates}
\NormalTok{plot\_est }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(data, }\AttributeTok{nperiod =} \DecValTok{6}\NormalTok{) \{}
\NormalTok{  joint\_data }\OtherTok{\textless{}{-}} \FunctionTok{do.call}\NormalTok{(ts.union, data)}
\NormalTok{  joint\_data }\OtherTok{\textless{}{-}}
    \FunctionTok{window}\NormalTok{(joint\_data,}
           \AttributeTok{start =}\NormalTok{ last\_dates[}\DecValTok{1}\NormalTok{] }\SpecialCharTok{{-}}\NormalTok{ nperiod }\SpecialCharTok{*} \FunctionTok{deltat}\NormalTok{(joint\_data))}

\NormalTok{  data\_legend }\OtherTok{\textless{}{-}}
    \FunctionTok{data.frame}\NormalTok{(}\AttributeTok{x =}\NormalTok{ last\_dates,}
               \AttributeTok{y =} \FunctionTok{sapply}\NormalTok{(data, tail, }\DecValTok{1}\NormalTok{),}
               \AttributeTok{label =} \FunctionTok{colnames}\NormalTok{(joint\_data))}

\NormalTok{  forecast}\SpecialCharTok{::}\FunctionTok{autoplot}\NormalTok{(joint\_data) }\SpecialCharTok{+} \FunctionTok{theme\_bw}\NormalTok{() }\SpecialCharTok{+}
    \FunctionTok{scale\_x\_continuous}\NormalTok{(}\AttributeTok{labels =}\NormalTok{ zoo}\SpecialCharTok{::}\NormalTok{as.yearmon) }\SpecialCharTok{+}
    \FunctionTok{geom\_text}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ x, }\AttributeTok{y =}\NormalTok{ y, }\AttributeTok{label =}\NormalTok{ label, }\AttributeTok{colour =}\NormalTok{ label),}
              \AttributeTok{data =}\NormalTok{ data\_legend,}
              \AttributeTok{check\_overlap =} \ConstantTok{TRUE}\NormalTok{, }\AttributeTok{hjust =} \DecValTok{0}\NormalTok{, }\AttributeTok{nudge\_x =} \FloatTok{0.01}\NormalTok{,}
              \AttributeTok{size =} \DecValTok{2}\NormalTok{, }\AttributeTok{inherit.aes =} \ConstantTok{FALSE}\NormalTok{) }\SpecialCharTok{+}
    \FunctionTok{theme}\NormalTok{(}\AttributeTok{legend.position =} \StringTok{"none"}\NormalTok{)  }\SpecialCharTok{+}
    \FunctionTok{labs}\NormalTok{(}\AttributeTok{x =} \ConstantTok{NULL}\NormalTok{, }\AttributeTok{y =} \ConstantTok{NULL}\NormalTok{)}
\NormalTok{\}}

\NormalTok{plot\_prevs }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{ (data, }\AttributeTok{nperiod =} \DecValTok{6}\NormalTok{) \{}
\NormalTok{  joint\_data }\OtherTok{\textless{}{-}} \FunctionTok{do.call}\NormalTok{(ts.union, }\FunctionTok{lapply}\NormalTok{(data, }\ControlFlowTok{function}\NormalTok{(x) \{}
\NormalTok{    first\_date }\OtherTok{\textless{}{-}} \FunctionTok{time}\NormalTok{(x)[}\DecValTok{1}\NormalTok{] }\SpecialCharTok{{-}} \FunctionTok{deltat}\NormalTok{(x)}
    \CommentTok{\# The last observed data is added for readability}
    \FunctionTok{ts}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\FunctionTok{window}\NormalTok{(y, }\AttributeTok{start =}\NormalTok{ first\_date, }\AttributeTok{end =}\NormalTok{ first\_date), x),}
       \AttributeTok{start =}\NormalTok{ first\_date, }\AttributeTok{frequency =} \FunctionTok{frequency}\NormalTok{(x))}
\NormalTok{  \}))}

\NormalTok{  data\_legend }\OtherTok{\textless{}{-}}
    \FunctionTok{data.frame}\NormalTok{(}\AttributeTok{x =} \FunctionTok{sapply}\NormalTok{(data, }\ControlFlowTok{function}\NormalTok{(x) }\FunctionTok{tail}\NormalTok{(}\FunctionTok{time}\NormalTok{(x), }\DecValTok{1}\NormalTok{)),}
               \AttributeTok{y =} \FunctionTok{sapply}\NormalTok{(data, tail, }\DecValTok{1}\NormalTok{),}
               \AttributeTok{label =} \FunctionTok{colnames}\NormalTok{(joint\_data))}
\NormalTok{  forecast}\SpecialCharTok{::}\FunctionTok{autoplot}\NormalTok{(joint\_data, }\AttributeTok{linetype =} \StringTok{"dashed"}\NormalTok{) }\SpecialCharTok{+}
\NormalTok{    forecast}\SpecialCharTok{::}\FunctionTok{autolayer}\NormalTok{(}
      \FunctionTok{window}\NormalTok{(y, }\AttributeTok{start =}\NormalTok{ last\_dates[}\DecValTok{1}\NormalTok{] }\SpecialCharTok{{-}}\NormalTok{ nperiod }\SpecialCharTok{*} \FunctionTok{deltat}\NormalTok{(y)),}
      \AttributeTok{colour =} \ConstantTok{FALSE}
\NormalTok{    ) }\SpecialCharTok{+}
    \FunctionTok{theme\_bw}\NormalTok{() }\SpecialCharTok{+}
    \FunctionTok{scale\_x\_continuous}\NormalTok{(}\AttributeTok{labels =}\NormalTok{ zoo}\SpecialCharTok{::}\NormalTok{as.yearmon) }\SpecialCharTok{+}
    \FunctionTok{geom\_text}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ x, }\AttributeTok{y =}\NormalTok{ y, }\AttributeTok{label =}\NormalTok{ label, }\AttributeTok{colour =}\NormalTok{ label),}
              \AttributeTok{data =}\NormalTok{ data\_legend,}
              \AttributeTok{check\_overlap =} \ConstantTok{TRUE}\NormalTok{, }\AttributeTok{hjust =} \DecValTok{0}\NormalTok{, }\AttributeTok{nudge\_x =} \FloatTok{0.01}\NormalTok{,}
              \AttributeTok{size =} \DecValTok{2}\NormalTok{, }\AttributeTok{inherit.aes =} \ConstantTok{FALSE}\NormalTok{) }\SpecialCharTok{+}
    \FunctionTok{theme}\NormalTok{(}\AttributeTok{legend.position =} \StringTok{"none"}\NormalTok{) }\SpecialCharTok{+}
    \FunctionTok{labs}\NormalTok{(}\AttributeTok{x =} \ConstantTok{NULL}\NormalTok{, }\AttributeTok{y =} \ConstantTok{NULL}\NormalTok{)}
\NormalTok{\}}

\NormalTok{all\_est }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(lp\_est, }\FunctionTok{list}\NormalTok{(}\StringTok{"LC local param."} \OtherTok{=}\NormalTok{ loc\_lc\_est),}
             \FunctionTok{list}\NormalTok{(}\StringTok{"QL local param."} \OtherTok{=}\NormalTok{ loc\_ql\_est))}
\NormalTok{all\_if }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(lp\_if, }\FunctionTok{list}\NormalTok{(}\StringTok{"LC local param."} \OtherTok{=}\NormalTok{ loc\_lc\_if),}
            \FunctionTok{list}\NormalTok{(}\StringTok{"QL local param."} \OtherTok{=}\NormalTok{ loc\_ql\_if))}
\NormalTok{y\_lim }\OtherTok{\textless{}{-}} \ConstantTok{NULL}
\NormalTok{all\_plots\_est }\OtherTok{\textless{}{-}} \FunctionTok{lapply}\NormalTok{(}
  \FunctionTok{names}\NormalTok{(all\_est),}
  \ControlFlowTok{function}\NormalTok{(x) }\FunctionTok{plot\_est}\NormalTok{(all\_est[[x]]) }\SpecialCharTok{+}
    \FunctionTok{ggtitle}\NormalTok{(latex2exp}\SpecialCharTok{::}\FunctionTok{TeX}\NormalTok{(}\FunctionTok{sprintf}\NormalTok{(}\StringTok{"Trend{-}cycle with \%s"}\NormalTok{, x))) }\SpecialCharTok{+}
    \FunctionTok{coord\_cartesian}\NormalTok{(}\AttributeTok{ylim =}\NormalTok{ y\_lim)}
\NormalTok{)}
\NormalTok{all\_plots\_prev }\OtherTok{\textless{}{-}} \FunctionTok{lapply}\NormalTok{(}
  \FunctionTok{names}\NormalTok{(all\_if),}
  \ControlFlowTok{function}\NormalTok{(x) }\FunctionTok{plot\_prevs}\NormalTok{(all\_if[[x]]) }\SpecialCharTok{+}
    \FunctionTok{ggtitle}\NormalTok{(latex2exp}\SpecialCharTok{::}\FunctionTok{TeX}\NormalTok{(}\FunctionTok{sprintf}\NormalTok{(}\StringTok{"Implicit forecasts with \%s"}\NormalTok{, x)))}
\NormalTok{)}
\CommentTok{\# Combine all plots:}
\FunctionTok{wrap\_plots}\NormalTok{(all\_plots\_est, }\AttributeTok{ncol =} \DecValTok{2}\NormalTok{)}
\FunctionTok{wrap\_plots}\NormalTok{(all\_plots\_prev, }\AttributeTok{ncol =} \DecValTok{2}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\end{document}