### 2024-02-12 supp_material.qmd

---
title: "Supplemental material"
format:
  pdf:
    documentclass: article
    number-sections: true
    pdf-engine: pdflatex
    toc: false
    include-in-header: [preamble.tex]
    #include-before-body: template/preambule.tex
    geometry:
      - top=2cm
      - bottom=2cm
      - left = 2.5cm
      - right = 2.5cm
linestretch: 2
echo: false
tbl-cap-location: top
fig-cap-location: top
---

\appendix

# Coefficients, gain and phase shift functions {#sec-an-cof}

```{r, out.width="100%"}
#| label: fig-graphs-coef-lc
#| fig-cap: Coefficients, gain and phase shift functions for the Linear-Constant (LC) filter with $I/C=3.5$.
img <- sprintf("img/filters_used/%s.%s", tolower("LC"), "pdf")
knitr::include_graphics(img)
```

```{r, out.width="100%"}
#| label: fig-graphs-coef-ql
#| fig-cap: Coefficients, gain and phase shift functions for the Quadratic-Linear (QL) filter with $I/C=3.5$.
img <- sprintf("img/filters_used/%s.%s", tolower("QL"), "pdf")
knitr::include_graphics(img)
```

```{r, out.width="100%"}
#| label: fig-graphs-coef-cq
#| fig-cap: Coefficients, gain and phase shift functions for the Cubic-Quadratic (CQ) filter with $I/C=3.5$.
img <- sprintf("img/filters_used/%s.%s", tolower("CQ"), "pdf")
knitr::include_graphics(img)
```

```{r, out.width="100%"}
#| label: fig-graphs-coef-daf
#| fig-cap: Coefficients, gain and phase shift functions for the direct asymmetric filter (DAF).
img <- sprintf("img/filters_used/%s.%s", tolower("DAF"), "pdf")
knitr::include_graphics(img)
```

\newpage

# R code example {#ann-ex-r}

In this appendix we present the {{< fa brands r-project >}} code that can be used to estimate trend-cycle components for the US Employment with local polynomial filters.

```{r, eval=FALSE, echo=TRUE}
#| eval: false
#| echo: true
#| file: R_ex.R
```