### 2024-04-02 paper/supp_material.qmd
paer folder

---
title: "Supplemental material"
format:
  pdf:
    documentclass: article
    number-sections: true
    pdf-engine: pdflatex
    toc: false
    include-in-header: [preamble.tex]
    #include-before-body: template/preambule.tex
    geometry:
      - top=2cm
      - bottom=2cm
      - left = 2.5cm
      - right = 2.5cm
linestretch: 2
echo: false
tbl-cap-location: top
fig-cap-location: top
---

\appendix

# Coefficients, gain and phase shift functions {#sec-an-cof}

```{r, out.width="100%"}
#| label: fig-graphs-coef-lc
#| fig-cap: Coefficients, gain and phase shift functions for the Linear-Constant (LC) filter with $I/C=3.5$.
img <- sprintf("img/filters_used/%s.%s", tolower("LC"), "pdf")
knitr::include_graphics(img)
```

```{r, out.width="100%"}
#| label: fig-graphs-coef-ql
#| fig-cap: Coefficients, gain and phase shift functions for the Quadratic-Linear (QL) filter with $I/C=3.5$.
img <- sprintf("img/filters_used/%s.%s", tolower("QL"), "pdf")
knitr::include_graphics(img)
```

```{r, out.width="100%"}
#| label: fig-graphs-coef-cq
#| fig-cap: Coefficients, gain and phase shift functions for the Cubic-Quadratic (CQ) filter with $I/C=3.5$.
img <- sprintf("img/filters_used/%s.%s", tolower("CQ"), "pdf")
knitr::include_graphics(img)
```

```{r, out.width="100%"}
#| label: fig-graphs-coef-daf
#| fig-cap: Coefficients, gain and phase shift functions for the direct asymmetric filter (DAF).
img <- sprintf("img/filters_used/%s.%s", tolower("DAF"), "pdf")
knitr::include_graphics(img)
```

```{r, out.width="100%"}
#| label: fig-graphs-coef-lc-nn
#| fig-cap: Coefficients, gain and phase shift functions for the Linear-Constant (LC) filter with $I/C=3.5$ and a nearest neighbour bandwidth.
img <- sprintf("img/filters_used/%s_nn.%s", tolower("LC"), "pdf")
knitr::include_graphics(img)
```

```{r, out.width="100%"}
#| label: fig-graphs-coef-ql-nn
#| fig-cap: Coefficients, gain and phase shift functions for the Quadratic-Linear (QL) filter with $I/C=3.5$ and a nearest neighbour bandwidth.
img <- sprintf("img/filters_used/%s_nn.%s", tolower("QL"), "pdf")
knitr::include_graphics(img)
```

```{r, out.width="100%"}
#| label: fig-graphs-coef-cq-nn
#| fig-cap: Coefficients, gain and phase shift functions for the Cubic-Quadratic (CQ) filter with $I/C=3.5$ and a nearest neighbour bandwidth.
img <- sprintf("img/filters_used/%s_nn.%s", tolower("CQ"), "pdf")
knitr::include_graphics(img)
```

```{r, out.width="100%"}
#| label: fig-graphs-coef-daf-nn
#| fig-cap: Coefficients, gain and phase shift functions for the direct asymmetric filter (DAF) with a nearest neighbour bandwidth.
img <- sprintf("img/filters_used/%s_nn.%s", tolower("DAF"), "pdf")
knitr::include_graphics(img)
```

\newpage

<!-- # R code example {#ann-ex-r} -->

<!-- In this appendix we present the {{< fa brands r-project >}} code that can be used to estimate trend-cycle components for the US Employment with local polynomial filters. -->

<!-- ```{r, eval=FALSE, echo=TRUE} -->
<!-- #| eval: false -->
<!-- #| echo: true -->
<!-- #| file: R_ex.R -->
<!-- ``` -->