### 2024-04-25 paper/article_jos.tex
update article

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
]{article}

\usepackage{amsmath,amssymb}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[top=2cm,bottom=2cm,left = 2.5cm,right = 2.5cm]{geometry}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage{stmaryrd}
\usepackage{multicol}
\usepackage{graphicx}
\usepackage{animate, dsfont, here, xspace}
%\usepackage{tikz}
\usepackage{tikz,pgfplots}
 \pgfplotsset{compat=1.17}
%\includepdf[fitpaper=true, pages=-]{img/pdg.pdf}

\DeclareMathOperator{\e}{e}
\DeclareMathOperator{\Determinant}{det}
\renewcommand{\P}{\mathds{P}} %Apparement \P existe déjà ?
\newcommand\N{\mathds{N}}
\newcommand\Norm{\mathcal{N}}
\newcommand\R{\mathds{R}}
\newcommand\Z{\mathds{Z}}
\newcommand\transp[1]{{#1}^T}
%\newcommand\C{\mathds{C}}
%\newcommand\Z{\mathds{Z}}

\newcommand\1{\mathds{1}}
\newcommand{\E}[2][]{{\mathds{E}}_{#1}
  \def\temp{#2}\ifx\temp\empty
  \else
    \left[#2\right]
  \fi
}
\newcommand{\V}[2][]{{\mathds{V}}_{#1}
  \def\temp{#2}\ifx\temp\empty
  \else
    \left[#2\right]
  \fi
}
\newcommand\ud{\,\mathrm{d}}
\newcommand{\ps}[2]{\left\langle #1 \,,\, #2 \right\rangle}

% blocks
\usepackage{environ}
\usepackage[tikz]{bclogo}

\tikzstyle{titlestyle} =[draw=black!80,fill=black!20, text=black,
 right=10pt, rounded corners]
\mdfdefinestyle{symmaryboxstyle}{
	linecolor=black!80, backgroundcolor = black!5,
	skipabove=\baselineskip, innertopmargin=\baselineskip,
	innerbottommargin=\baselineskip,
	userdefinedwidth=\textwidth,
	middlelinewidth=1.2pt, roundcorner=5pt,
	skipabove={\dimexpr0.5\baselineskip+\topskip\relax},
	frametitleaboveskip=\dimexpr-\ht\strutbox\relax,
	innerlinewidth=0pt,
}
\newcounter{summarybox}
\NewEnviron{summary_box}[2][true]{%
\refstepcounter{summarybox} % on incrémente le compteur
\begin{mdframed}[style=symmaryboxstyle,
nobreak=#1,
frametitle={%
      \tikz[baseline=(current bounding box.east),outer sep=0pt]
      \node[titlestyle,anchor=east]
    {Encadré \thesummarybox{} --- #2};}
]
\vspace{-0.5em}
\BODY
\end{mdframed}
}

\usepackage{amsthm}
%\theoremstyle{remark}
\newtheorem*{remarque}{Remarque}

\usepackage{mathrsfs}
\usepackage{fontawesome5}

%\usepackage[style=authoryear,%,uniquename=false, uniquelist=false,
%maxbibnames=2]{biblatex-chicago}
% \usepackage[options]{natbib}
% \bibliographystyle{chicago}

%\DefineBibliographyStrings{english}{andothers={et\addabbrvspace alii}}
\usepackage[authordate16,backend=biber,maxcitenames=2]{biblatex-chicago}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{biblatex}
\addbibresource{biblio.bib}
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Improving real-time trend estimates using local parametrisation of polynomial regression filters},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Improving real-time trend estimates using local parametrisation
of polynomial regression filters}
\author{}
\date{}

\begin{document}
\maketitle

\setstretch{2}
\section*{Abstract}\label{abstract}
\addcontentsline{toc}{section}{Abstract}

This paper studies and compares real-time estimates of the trend-cycle
component with moving averages based on local polynomial regression. In
particular, this theory allows us to reproduce Henderson's symmetric and
Musgrave's asymmetric filters used in the X-13ARIMA-SEATS seasonal
adjustment algorithm. This paper proposes two extensions of local
polynomial filters for real-time trend-cycle estimates (for the last
estimates): first including a timeliness criterion to minimise the phase
shift; second, with procedure for parametrising asymmetric filters
locally While they are generally parametrised globally, which can be
suboptimal around turning points. An empirical comparison, based on
simulated and real data, shows that modelling polynomial trends that are
too complex introduces more revisions without reducing the phase shift,
and that local parametrisation reduces the delay in detecting turning
points and reduces revisions. The paper also shows how these results can
be easily reproduced using the R package \texttt{rjd3filters} dedicated
to the manipulation of moving averages.

Keywords: time series, trend-cycle, seasonal adjustment, turning points,
R statistical software.

JEL Classification: E32, E37.

\section{Introduction}\label{introduction}

Analysis of the economic cycle, and in particular the early detection of
turning points in a series, is a major topic in the analysis of economic
outlook. To this end, economic indicators are generally seasonally
adjusted. However, in order to improve their interpretability, it may be
necessary to perform additional smoothing in order to reduce noise, and
thus analyse the trend-cycle component. By construction, trend-cycle
extraction methods are closely related to seasonal adjustment methods,
since they are generally applied to seasonally adjusted series.

A moving average, or linear filter, is a statistical method that
consists of applying a rolling weighted mean to a times series: for each
date \(t\) it computes a weighted mean of \(p\) past points and \(q\)
future points where \(p,q\geq0\) depends on the moving average. Moving
averages are ubiquitous in business cycle extraction and seasonal
adjustment methods. For example, the X-11 decomposition method
\autocite{x11}, used in X-13ARIMA-SEATS software \autocite{x13}, uses
several moving averages to estimate the main components of a time
series. Symmetric filters are applied to the center of the series, but
when it comes to estimate the most recent points, all of these methods
must rely on asymmetric filters. For trend-cycle extraction, the most
popular symmetric filter is the \textcite{henderson1916note} moving
average, which is used in the X-11 seasonal adjustment algorithm.

However, for real-time estimates, due to the lack of future
observations, all these methods must rely on asymmetric filters to
estimate the most recent points. For example, for trend-cycle
extraction, X-11 uses Henderson's symmetric filter and
\textcite{musgrave1964set}'s asymmetric filters on an extended series
using an ARIMA model. Since the predicted values are linear combinations
of past values, this amounts to applying asymmetric averages at the end
of the series.

If the classic asymmetric moving averages have good properties regarding
the future revisions induced by the process (see for example
\textcite{pierce1980SA}), they create, by construction, phase shifts
that impact the real-time estimation of turning points, introducing time
delay in the detection.

Using local polynomial regression, \textcite{proietti2008} developed a
general class of symmetric and asymmetric moving averages. In
particular, it becomes possible to reproduce Henderson's symmetric
filter and Musgrave's asymmetric filters. The aim of this study is to
propose two extensions to the \textcite{proietti2008} class of filters
in order to two drawbacks of those methods. Firstly, adding, in the
optimisation process, a timeliness criterin to directly control the
phase shift (i.e.~the delay in detecting turning points). Secondly,
calibrating asymmetric filter locally, while they are generally
calibrated globally using thus assumptions that may be wrong locally.
Those extensions are implemented in the statistical software R
\autocite{logicielr} throw the \texttt{rjd3filters} package (available
at \url{https://github.com/rjdemetra/rjd3filters}) and can therefore be
easily used. All the code associated to this paper is available at
\url{https://github.com/XXXX} (anonymized linked before the manuscript
is accepted).

In section~\ref{sec-propMM}, we describe the general properties of
moving averages and the associated quality criteria. This allows us to
understand the foundations behind the construction of moving averages
from local polynomial regressions, as well as those behind the two
extensions proposed in this article (section~\ref{sec-nonparamreg}). In
section~\ref{sec-comparison}, all these methods are compared empirically
on simulated and real series.

\section{Local trend-cycle models and moving averages}\label{sec-propMM}

The basic assumption used in time series decomposition methods is that
the input observed time series, \(y_t\) can be decomposed into a signal
component \(\mu_t\) and a an erratic component \(\varepsilon_t\) (called
irregular): \[
f(y_t)=\mu_t+\varepsilon_t
\] where \(f\) denotes some appropriate transformation (usually
logarithmic or no transformation). To simplify further notations,
\(y_t\) will denote the transformed observed series. The noise component
\(\varepsilon_t\) is usually assumed to be a white noise. Assuming that
the input time series is seasonally adjusted (or without seasonality),
the signal \(\mu_t\) represents the trend and cycle components, here
estimated jointly and called trend-cycle \(TC_t\). This component can be
locally approximated by a polynomial degree (see
section~\ref{sec-nonparamreg}): \[
\mu_t = \sum_{j=0}^d\beta_jt^j+\xi_t
\] with \(\xi_t\) a stochastic process uncorrelated with
\(\varepsilon_t\). Even if some papers model \(\xi_t\) and
\(\varepsilon_t\) separately \autocite[see for
example][]{GrayThomson2002}, an usual assumption, adopted in this paper,
is to is to gather \(\xi_t\) and \(\varepsilon_t\). Thus, the
trend-cycle \(TC_t\) is considered as deterministic and modelled as a
polynomial trend of degree \(d\). The coefficients
\((\beta_0,\dots,\beta_k)\) can be estimated by weighted least squares.
The estimate \(\hat \beta_0\) provides the trend-cycle estimate
\(\hat TC_t\) and it can be shown that this equivalent to apply of
moving average (see section~\ref{sec-nonparamreg}). The goal of this
paper is to present two new methods to separate the trend-cycle
component from the noise component in real-time (i.e.~when no future
point is known), using moving averages build on local polynomial
smoothing techniques.

Numerous papers describe the definition and the properties of moving
averages and linear filters, see for example \textcite{ch12HBSA}. In
this section we summarize some of the main results to understand the
next sections.

Let \(p\) and \(f\) two integers, a moving average
\(M_{\boldsymbol\theta}\) or \(M\) is defined by a set of coefficients
\(\boldsymbol\theta=\transp{\begin{pmatrix}\theta_{-p}&\cdots&\theta_{f}\end{pmatrix}}\)
such as for all time series \(y_t\): \[
M_{\boldsymbol\theta}(y_t)=\sum_{k=-p}^{+f}\theta_kX_{t+k}.
\]

\begin{itemize}
\item
  \(p+f+1\) is called the \emph{moving average order}.
\item
  When \(p=f\) the moving average is said to be \emph{centered}. If we
  also have \(\forall k:\:\theta_{-k} = \theta_k\), the moving average
  \(M_{\boldsymbol\theta}\) is said to be \emph{symmetric}. In this
  case, the quantity \(h=p=f\) is called the \emph{bandwidth}.
\end{itemize}

\subsection{Gain and phase shift functions}\label{subsec:gain-deph}

To interpret the notions of gain and phase shift, it is useful to
illustrate the effects of moving averages on harmonic series
\(y_t=\e^{-i\omega t}\) with \(\omega\in[0,\pi]\). The moving average
\(M_{\boldsymbol\theta}\) transforms \(y_t\) into: \[
Y_t = M_{\boldsymbol\theta}y_t = \sum_{k=-p}^{+f} \theta_k \e^{-i \omega (t+k)}
= \left(\sum_{k=-p}^{+f} \theta_k \e^{-i \omega k}\right)\cdot y_t.
\] The function
\(\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k e^{-i \omega k}\)
is called the \emph{transfer function} or \emph{frequency response
function}. The frequency response function can alternatively be defined
as
\(\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k e^{i \omega k}\)
or
\(\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k e^{2\pi i \omega k}\).
It can be rewritten as: \[
\Gamma_{\boldsymbol\theta}(\omega) = \rho_{\boldsymbol\theta}(\omega)\e^{i\varphi_{\boldsymbol\theta}(\omega)},
\] where
\(\rho_{\boldsymbol\theta}(\omega)=G_{\boldsymbol\theta}(\omega)=\lvert\Gamma_{\boldsymbol\theta}(\omega)\rvert\)
is the \emph{gain} or \emph{amplitude} function and
\(\varphi_{\boldsymbol\theta}(\omega)\) is the \emph{phase shift} or
\emph{time shift} function. We this have: \begin{align*}
\rho_{\boldsymbol\theta}(\omega) &=\sqrt{\left(\sum_{k=-p}^{+f}\theta_k\cos(\omega k)\right)^2+
\left(\sum_{k=-p}^{+f}\theta_k\sin(\omega k)\right)^2} \\
\tan \varphi_{\boldsymbol\theta}(\omega) &=- \frac{\sum_{k=-p}^{+f}\theta_k\sin(\omega k)}{\sum_{k=-p}^{+f}\theta_k\cos(\omega k)} \text{ if }\sum_{k=-p}^{+f}\theta_k\cos(\omega k) \ne 0 \\
&= 0 \text{ if }\sum_{k=-p}^{+f}\theta_k\cos(\omega k) = 0.
\end{align*} The phase shift is sometimes represented as
\(\phi_{\boldsymbol\theta}(\omega)=\frac{\varphi_{\boldsymbol\theta}(\omega)}{\omega}\)
to measure it in number of periods. For all symmetric moving averages we
have \(\varphi_{\boldsymbol\theta}(\omega)\equiv 0 \;(modulo\;{\pi})\).

To sum up, applying a moving average to a harmonic times series
(\(y_t=\e^{-i\omega t}\)) affects it in in two different ways:

\begin{itemize}
\item
  by multiplying it by an amplitude coefficient
  \(\rho_{\boldsymbol\theta}\left(\omega\right)\) (gain);
\item
  by ``shifting'' it in time by
  \(\varphi_{\boldsymbol\theta}(\omega)/\omega\) which directly affects
  the detection of turning points. When
  \(\varphi_{\boldsymbol\theta}(\omega)/\omega<0\) the time shift is
  negative: the turning point is detected with delay.
\end{itemize}

Fourier decomposition allows us to analyze any time series as a sum of
harmonic series, and each component (trend, cycle, seasonal, irregular)
is associated with a set of frequencies. Noting \(\omega = 2\pi/s\), the
harmonic series of frequency \(\omega\) represents a series that repeats
itself every \(s\) periods. For example, for a monthly series (\(s=12\)
observations per year), seasonal movements are those that repeat every
year: they are therefore associated with frequencies \(2\pi/12\) (annual
periodicity), \(2\pi/12\times 2=2\pi/6,\dots,2\pi/12\times 5\). In this
paper, we consider that the trend-cycle is associated with frequencies
in the interval \([0, 2\pi/12]\), i.e.~movements recurring at least
every 12 months. Even if different frequencies are sometimes used
(e.g.~\([0, 2\pi/36]\) to consider only cycles of at least 36 months),
this has no impact on the results of the study. The other frequencies,
\([2\pi/12, \pi]\) are associated with the irregular component.

Figure~\ref{fig-graphsmusgrave} shows the gain and phase shift function
for the asymmetric Musgrave filter (see section~\ref{sec-lppasymf})
often used for real-time trend-cycle estimation (i.e.~when no point in
the future is known). The gain function is greater than 1 on the
frequencies associated with the trend-cycle (\([0, 2\pi/12]\)): this
means that the trend-cycle is well preserved and that short cycles of 1
to 2 years (\([2\pi/24,2\pi/12]\)) are even amplified. At other
frequencies, the gain function is less than 1, but always positive: this
means that the series smoothed by this moving average will always
contain noise, even if it is attenuated. Analysis of the phase shift
shows that the shorter the cycles, the greater the phase shift: this
means that on series smoothed by this moving average, turning points
could be detected at the wrong date.

\begin{figure}

\caption{\label{fig-graphsmusgrave}Coefficients, gain and phase shift
function for the Musgrave filter used in X-11 in real time (last
estimate) for monthly series when the bandwidth is 6.}

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{img/filters_used/musgrave.pdf}

Note: the Musgrave filter depends on a ratio \(I/C\) (see
section~\ref{sec-nonparamreg}). In X-11, for monthly series, when the
bandwidth is 6 (i.e.: a symmetric Henderson filter of order 13 is used,
what is used in most cases), the ratio is fixed to 3.5 and this value is
used in this figure.

}

\end{figure}%

\subsection{Desirable properties of a moving
average}\label{desirable-properties-of-a-moving-average}

To decompose a time series into a seasonal component, a trend-cycle and
the irregular, the X-11 decomposition algorithm (used in
X-13ARIMA-SEATS) uses a succession of moving averages, all with specific
constraints.

In this paper, we assume that our initial series \(y_t\) is seasonally
adjusted and can be written as the sum of a trend-cycle, \(TC_t\), and
an irregular (noise) component, \(\varepsilon_t\): \[
y_t=TC_t+\varepsilon_t.
\] The aim will be to build moving averages that best preserve the
trend-cycle (\(M_{\boldsymbol\theta} (TC_t)\simeq TC_t\)).

Trend-cycles are generally modelled by local polynomial trends (see
section~\ref{sec-nonparamreg}) and, in order to best preserve the
trend-cycles, we thus design moving averages that preserve the
polynomial trends. A moving average \(M_{\boldsymbol\theta}\) preserves
a function of time \(f(t)\) if
\(\forall t:M_{\boldsymbol\theta} f(t)=f(t)\).

For the moving average \(M_{\boldsymbol\theta}\), it can be shown that:

\begin{itemize}
\tightlist
\item
  \(M_{\boldsymbol\theta}\) preserves polynomials of degree \(d\) if and
  only if: \[
  \sum_{k=-p}^{+f}\theta_k=1
   \text{ and }
  \forall j \in \left\{1,2,\dots,d\right\}:\:
  \sum_{k=-p}^{+f}k^j\theta_k=0.
  \]
\item
  If \(M_{\boldsymbol\theta}\) is symmetric (\(p=f\) and
  \(\theta_{-k} = \theta_k\)) and preserves polynomials of degree \(2d\)
  then it also preserves polynomials of degree \(2d+1\).
\end{itemize}

\subsection{Real-time estimation and asymmetric moving
average}\label{sec-mmasym}

For symmetric filters, the phase shift function is equal to zero (modulo
\(\pi\)). So there is no delay in detecting turning points. However,
they cannot be used in the beginning and in the end of the time series
because no past or future value can be used. Thus, for real-time
estimation, it is needed to build asymmetric moving averages that
approximate the symmetric moving average.

Several approaches can be used for real-time estimation:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Use asymmetric moving averages to take account of the lack of
  available data;
\item
  Apply symmetric filters to extended forecast series. This method seems
  to date back to \textcite{deforest1877adjustment}, which also suggests
  modelling a polynomial trend of degree three or less at the end of the
  period. A similar approach is used in the X-13ARIMA-SEATS seasonal
  adjustment method: the series is extend by one year (by default) with
  an ARIMA model to minimise the revisions associated to asymmetric
  filters. Since the final filter used for the seasonal adjustment
  extraction is much longer than one year (in most of the cases, for
  monthly series, the bandwidth is 78), the input series is just
  partially extended (and asymmetric filters are still used) to prevent
  from instability over long periods (e.g., outliers or model shifts).
\end{enumerate}

Ultimately, the second method is equivalent to using asymmetric moving
averages, since forecasts are linear combinations of the past.

Conversely, the implicit forecasts of an asymmetric moving average can
be deduced from a reference symmetric moving average. This allows us to
judge the quality of real-time estimates of the trend-cycle and to
anticipate future revisions when forecasts are far from expected
evolutions.\\
Let us denote \(\boldsymbol v=(v_{-h},\dots, v_{h})\) the reference
symmetric moving average (used to compute the final estimates) and
\(\boldsymbol w^0,\dots, \boldsymbol w^{h-1}\) a sequence of asymmetric
moving averages used for the intermediate estimates.
\(\boldsymbol w^0=(w_{-h}^0, \dots, w_{0}^0)\) is used for estimation in
real time estimate (when 0 point in the future is known),
\(\boldsymbol w^1=(w_{-h}^1, \dots, w_{0}^1, w_{1}^1)\) is used for
estimation of the penultimate point (when only 1 in the future is
known), and so on. Let \(y_{t-h},\dots,y_{t}\) be the observed series
during the last \(h\) periods (for which asymmetric moving averages are
used to get the intermediate trend-cycle estimates for the dates \(t-h\)
to \(h\)). The implicit forecast \(y_{t+1}^*,\dots y_{t+h}^*\) induced
by \(w^0,\dots w^{h-1}\) are the forecasts of the series \(y_t\) for
which the estimates of the symmetric filter applied to the extended
series
\(\boldsymbol y^* = (y_{t-h}, \dots, y_t, y_{t+1}^*,\dots y_{t+h}^*)\)
produce the same estimates as the asymmetric filter applied to the
series \(\boldsymbol y^*\). This means: \begin{align*}
\underbrace{\sum_{i=-h}^0 v_iy_{ti} + \sum_{i=1}^h v_iy_{t+i}^*}_{\text{smoothing by }\boldsymbol v\text{ of the extended series}} &=
\underbrace{\sum_{i=-h}^0 w_i^0y_{t+i}}_{\text{smoothing by }\boldsymbol w^0\text{ using }0\text{ points after }t} \\
&=
\underbrace{\sum_{i=-h}^0 w_i^1y_{t+i} + w_1^1y_{t+1}^*}_{\text{smoothing by }\boldsymbol w^1\text{ using }1\text{ points after }t} \\
&= \underbrace{\sum_{i=-h}^0 w_i^2y_{t+i} + \sum_{i=1}{2}w_i^2y_{t+i}^*}_{\text{smoothing by }\boldsymbol w^2\text{ using }2\text{ points after }t} \\
&= \cdots \\
&=\underbrace{\sum_{i=-h}^0 w_i^{h-1}y_{t+i} + \sum_{i=1}^{h-1} w_i^{h-1}y_{t+i}^*}_{\text{smoothing by }\boldsymbol w^{h-1}\text{ using }h-1\text{ points after }t}
\end{align*} For any integer \(q\) between 0 and \(h-1\), using for
convention \(w_t^q=0\) if \(t>q\) we then have: \[
\sum_{i=1}^h (v_i- w_i^q) y_i^*
=\sum_{i=-h}^0 (w_i^q-v_i)y_i.
\] In matrix form, this is equivalent to solving: \[
\scriptstyle
\begin{pmatrix}
  v_1 & v_2 & \cdots & v_h \\
  v_1 - w_1^1 & v_2 & \cdots & v_h \\
  \vdots & \vdots & \cdots & \vdots \\
   v_1 - w_1^{h-1} & v_2-w_2^{h-1} & \cdots & v_h
\end{pmatrix}
\begin{pmatrix}y_1^* \\ \vdots \\ y_h^*\end{pmatrix}=
\begin{pmatrix}
  w_{-h}^0 - v_{-h} & w_{-(h-1)}^0 - v_{-(h-1)} & \cdots & w_{0}^0 - v_{0} \\
  w_{-h}^1 - v_{-h} & w_{-(h-1)}^1 - v_{-(h-1)} & \cdots & w_{0}^1 - v_{0} \\
  \vdots & \vdots & \cdots & \vdots \\
  w_{-h}^{h-1} - v_{-h} & w_{-(h-1)}^{h-1} - v_{-(h-1)} & \cdots & w_{0}^{h-1} - v_{0}
\end{pmatrix}
\begin{pmatrix}y_{-h} \\ \vdots \\ y_0\end{pmatrix}.
\] This is implemented in the \texttt{rjd3filters::implicit\_forecast()}
function.

As highlighted by \textcite{wildischis2004}, extending the series
through forecasting with an ARIMA model is equivalent to calculating
asymmetric filters whose coefficients are optimised in relation to the
one-step ahead forecast. In other words, the aim is to minimise the
revisions between the first and last estimates (with the symmetric
filter). However, the phase shift induced by the asymmetric filters is
not controlled: we might prefer to have faster detection of turning
points and a larger revision rather than just minimising the revisions
between the first and last estimates. Furthermore, since the
coefficients of the symmetric filter (and therefore the weight
associated with distant forecasts) decrease slowly, we should also be
interested in the performance of multi-step ahead forecasting. This is
why it may be necessary to define alternative criteria for judging the
quality of asymmetric moving averages.

\section{Non-parametric regression and local polynomial
regression}\label{sec-nonparamreg}

Many trend-cycle extraction methods are based on non-parametric
regressions, which are particularly flexible because they do not assume
any predetermined dependency in the predictors. In practice, local
regressions can be used. More specifically, consider a set of points
\((x_t,y_t)_{1\leq t\leq n}\) and we here consider that \(x_t=t\) (no
external regressor is used). Non-parametric regression involves assuming
that there exists a function \(\mu\), to be estimated, such that
\(y_t=\mu(t)+\varepsilon_t\) with \(\varepsilon_t\) an error term.
According to Taylor's theorem, for any date \(t_0\), if \(\mu\) is
differentiable \(d\) times, then:
\begin{equation}\phantomsection\label{eq-taylor}{
\forall t \::\:\underbrace{\mu(t)}_{=TC_t} = \underbrace{\mu(t_0)}_{=TC_{t_0}} + \mu'(t_0)(t-t_0)+\dots +
\frac{\mu^{(d)}(t_0)}{d!}(t-t_0)^d+R_d(t),
}\end{equation} where \(R_d\) is a negligible residual term in the
neighbourhood of \(t_0\). In a neighbourhood \(h(t_0)\) around \(t_0\),
\(\mu\) can therefore be approximated by a polynomial of degree \(d\).
The quantity \(h(t_0)\) is called the \emph{bandwidth}. If
\(\varepsilon_t\) is white noise, we can estimate \(\mu(t_0)=TC_{t_0}\)
by least squares using the observations which are in
\(\left[t_0-h(t_0),t_0+h(t_0)\right]\). In practice, this means assuming
that the trend is locally polynomial. We also generally assume that the
bandwith is fixed over time: \(h(t_0)=h.\) This is also what is assumed
in the seasonal adjustment algorithm X-11 and what will be assumed in
this paper.

Various estimation methods can be used to derive symmetric and
asymmetric moving averages.\\
\textcite{GrayThomson1996} propose a complete statistical framework that
makes it possible, in particular, to model the error in approximating
the trend by local polynomials. However, as the specification of this
error is generally complex, simpler models may be preferred, such as
that of \textcite{proietti2008}.\\
Finally, \textcite{dagumbianconcini2008} propose a similar modelling of
the trend-cycle but using the theory of Hilbert spaces with reproducing
kernels for estimation, which has the particular advantage of
facilitating the calculation of different moving averages at different
time frequencies. See \textcite{inseeDTM202401} for a more detailed
comparison of the different recent methods for trend-cycle extraction
included in the \texttt{rjd3filters} package

In this paper we will focus on \textcite{proietti2008} approach.
Section~\ref{sec-sym-lp} and section~\ref{sec-lppasymf} describes their
approach to build symmetric moving averages (used for the final
estimates of the trend-cycle component) and asymmetric moving averages
(used for real-time estimates). This paper proposed two extensions of
their approach to build asymetric moving averages: a first one adding a
criteria to control the phase shift (section~\ref{sec-localic}), not
used in the simulation (section~\ref{sec-comparison}) but which could be
used in future studied; a second one using a local parameterisation of a
parameter usually parameterised globally
(section~\ref{sec-lptimeliness}), which is used in the simulation.

\subsection{Symmetric moving averages and local polynomial
regression}\label{sec-sym-lp}

Using \textcite{proietti2008}'s notation, we assume that our time series
\(y_t\) can be decomposed into \[
y_t=\mu_t+\varepsilon_t,
\] where \(\mu_t\) is the trend and
\(\varepsilon_{t}\overset{i.i.d}{\sim}\mathcal{N}(0,\sigma^{2})\) is the
noise (the series is therefore seasonally adjusted). Within a
neighbourhood \(h\) of \(t\), the local trend \(\mu_t\) is approximated
by a polynomial of degree \(d\), such that \(\mu_t\simeq m_{t}\) with:
\[
\forall j\in\left\{-h,-h+1,\dots,h\right\}:\:
y_{t+j}=m_{t+j}+\varepsilon_{t+j},\quad m_{t+j}=\sum_{i=0}^{d}\beta_{i}j^{i}.
\] The problem of trend extraction is equivalent to estimating
\(m_t=\beta_0\) (the constant in the previous formula).

In matrix notation: \begin{equation}\phantomsection\label{eq-lpp}{
\underbrace{\begin{pmatrix}y_{t-h}\\
y_{t-(h-1)}\\
\vdots\\
y_{t}\\
\vdots\\
y_{t+(h-1)}\\
y_{t+h}
\end{pmatrix}}_{\boldsymbol y}=\underbrace{\begin{pmatrix}1 & -h & h^{2} & \cdots & (-h)^{d}\\
1 & -(h-1) & (h-1)^{2} & \cdots & (-(h-1))^{d}\\
\vdots & \vdots & \vdots & \cdots & \vdots\\
1 & 0 & 0 & \cdots & 0\\
\vdots & \vdots & \vdots & \cdots & \vdots\\
1 & h-1 & (h-1)^{2} & \cdots & (h-1)^{d}\\
1 & h & h^{2} & \cdots & h^{d}
\end{pmatrix}}_{\boldsymbol X}\underbrace{\begin{pmatrix}\beta_{0}\\
\beta_{1}\\
\vdots\\
\vdots\\
\vdots\\
\vdots\\
\beta_{d}
\end{pmatrix}}_{\boldsymbol \beta}+\underbrace{\begin{pmatrix}\varepsilon_{t-h}\\
\varepsilon_{t-(h-1)}\\
\vdots\\
\varepsilon_{t}\\
\vdots\\
\varepsilon_{t+(h-1)}\\
\varepsilon_{t+h}
\end{pmatrix}}_{\boldsymbol \varepsilon}.
}\end{equation}

Since \(d+1\) parameters are estimated with \(\boldsymbol \beta\), we
need at least \(2h+1\) observations (thus \(2h\geq d\)) and the
estimation is made by weighted least squares (WLS), which is equivalent
to minimising the following objective function: \[
S(\hat{\beta}_{0},\dots,\hat{\beta}_{d})=\sum_{j=-h}^{h}\kappa_{j}(y_{t+j}-\hat{\beta}_{0}-\hat{\beta}_{1}j-\dots-\hat{\beta}_{d}j^{d})^{2}.
\] where \(\kappa_j\) is a set of weights called a \emph{kernel}. We
have \(\kappa_j\geq 0:\kappa_{-j}=\kappa_j\), and with
\(\boldsymbol K=diag(\kappa_{-h},\dots,\kappa_{h})\), the estimate of
\(\boldsymbol \beta\) can be written as
\(\hat{\boldsymbol\beta}=(\transp{\boldsymbol X}\boldsymbol K\boldsymbol X)^{-1}\transp{\boldsymbol X}\boldsymbol K\boldsymbol y.\)
With
\(\boldsymbol e_1=\transp{\begin{pmatrix}1 &0 &\cdots&0 \end{pmatrix}}\),
the estimate of the trend is:
\begin{equation}\phantomsection\label{eq-mmsym}{
\hat{m}_{t}=\boldsymbol e_{1}\hat{\boldsymbol \beta}=\transp{\boldsymbol \theta}\boldsymbol y=\sum_{j=-h}^{h}\theta_{j}y_{t-j}\text{ with }\boldsymbol \theta=\boldsymbol K\boldsymbol X(\transp{\boldsymbol X}\boldsymbol K\boldsymbol X)^{-1}\boldsymbol e_{1}.
}\end{equation} To conclude, the estimate of the trend \(\hat{m}_{t}\)
can be obtained applying the symmetric filter \(\boldsymbol \theta\) to
\(y_t\) (\(\boldsymbol \theta\) is symmetric due to the symmetry of the
kernel weights \(\kappa_j\)). Moreover, since
\(\transp{\boldsymbol X}\boldsymbol \theta=\boldsymbol e_{1}\) we have:
\[
\sum_{j=-h}^{h}\theta_{j}=1,\quad\forall r\in\left\{1,2,\dots,d\right\}:\sum_{j=-h}^{h}j^{r}\theta_{j}=0.
\] Hence, the filter \(\boldsymbol w\) preserves deterministic
polynomials of order \(d\).

Regarding parameter selection, the general consensus is that the choice
between different kernels is not crucial. See, for example,
\textcite{cleveland1996smoothing} or \textcite{Loader1999}. The only
desired constraints on the kernel are that it assigns greater weight to
the central estimation (\(\kappa_0\)) and decreases towards 0 as it
moves away from the central estimation. The uniform kernel should
therefore be avoided. It is then preferable to focus on two other
parameters:

\begin{itemize}
\item
  the degree of the polynomial, denoted by \(d\): if it is too small,
  there is a risk of biased estimates of the trend-cycle, and if it is
  too large, there is a risk of excessive variance in the estimates (due
  to over-adjustment);
\item
  the number of neighbors \(2h+1\) (or the window \(h\)): if it is too
  small, then there will be too little data for the estimates (resulting
  in high variance in the estimates), and if it is too large, the
  polynomial approximation will probably be wrong, leading to biased
  estimates.
\end{itemize}

In this paper we will use the Henderson kernel: \[
\kappa_{j}=\left[1-\frac{j^2}{(h+1)^2}\right]
\left[1-\frac{j^2}{(h+2)^2}\right]
\left[1-\frac{j^2}{(h+3)^2}\right].
\] However, several other kernels are available in \texttt{rjd3filters}.

\subsection{Asymmetric moving averages and local polynomial
regression}\label{sec-lppasymf}

As mentioned in section~\ref{sec-mmasym}, for real-time estimation,
several approaches can be used:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Apply symmetric filters to the series extended by forecasting
  \(\hat{y}_{n+l\mid n},l\in\left\{1,2,\dots,h\right\}\).
\item
  Build an asymmetric filter by local polynomial approximation on the
  available observations (\(y_{t}\) for
  \(t\in\left\{n-h,n-h+1,\dots,n\right\}\)).
\item
  Build asymmetric filters that minimise the mean squared error of
  revision under polynomial trend reproduction constraints.
\end{enumerate}

\textcite{proietti2008} show that the first two approaches are
equivalent when forecasts are made by polynomial extrapolation of degree
\(d\). They are also equivalent to the third approach under the same
constraints as those of the symmetric filter. The third method is called
\emph{direct asymmetric filtering} (DAF). This method is used for
real-time estimation in the STL seasonal adjustment method
(\emph{Seasonal-Trend decomposition based on Loess}, see
\textcite{cleveland90}). Although DAF filter estimates are unbiased,
this is at the cost of greater variance in the estimates.

To solve the problem of the variance of real-time filter estimates,
\textcite{proietti2008} propose a general method for constructing
asymmetric filters that allows a bias-variance trade-off. This is a
generalisation of \textcite{musgrave1964set}'s asymmetric filters (used
in the X-11 seasonal adjustment algorithm).

Rewriting equation~\ref{eq-lpp}:
\begin{equation}\phantomsection\label{eq-lpgeneralmodel}{
\boldsymbol y=\begin{pmatrix}\boldsymbol U &\boldsymbol Z\end{pmatrix}
\begin{pmatrix}\boldsymbol \gamma \\ \boldsymbol \delta\end{pmatrix}+\boldsymbol \varepsilon,\quad
\boldsymbol \varepsilon\sim\mathcal{N}(0,\boldsymbol D).
}\end{equation} where \([\boldsymbol U,\boldsymbol Z]\) is of full rank
and forms a subset of the columns of \(X.\) Defining the variance error
as \(\boldsymbol D = \sigma^2\boldsymbol K^{-1}\) (to directly take into
account the weights associated to the observations) and with
\([\boldsymbol U,\boldsymbol Z] = \boldsymbol X\) we thus find the model
associated to symmetric filters. The objective is to find an asymmetric
filter \(\boldsymbol v\) which uses \(q\) points in the future and
minimises the mean square error of revision (to the symmetric filter
\(\boldsymbol \theta\)) under certain (polynomial) constraints. The
matrices \(\boldsymbol y\), \(\boldsymbol U\), \(\boldsymbol Z\) and
\(\boldsymbol \theta\) are partitioned as follows: \[
\boldsymbol y = \begin{pmatrix} \boldsymbol y_p \\ \boldsymbol y_f \end{pmatrix}, \quad
\boldsymbol U = \begin{pmatrix} \boldsymbol U_p \\ \boldsymbol U_f \end{pmatrix}, \quad  
\boldsymbol Z = \begin{pmatrix} \boldsymbol Z_p \\ \boldsymbol Z_f \end{pmatrix}, \quad 
\boldsymbol \theta = \begin{pmatrix} \boldsymbol \theta_p \\ \boldsymbol \theta_f \end{pmatrix},
\] where \(\boldsymbol y_p\) is a matrix \((h+q+1)\times 1\) which
denotes the set of available observations (from \(y_{t-h}\) to
\(y_{t+q}\)) and \(\boldsymbol U\), \(\boldsymbol Z\) and
\(\boldsymbol \theta\) are partitioned in the same way. The constraints
are defined as
\(\transp{\boldsymbol U_p}\boldsymbol v=\transp{\boldsymbol U}\boldsymbol \theta\):
they represents the polynomial preservation constraints we want to
impose on the asymmetric filter. \textcite{proietti2008} show that the
problem is equivalent to finding \(v\) which minimises:
\begin{equation}\phantomsection\label{eq-lppasym}{
\varphi(\boldsymbol v)=
\underbrace{
  \underbrace{\transp{(\boldsymbol v-\boldsymbol \theta_{p})}\boldsymbol D_{p}(\boldsymbol v-\boldsymbol \theta_{p})+
  \transp{\boldsymbol \theta_{f}}\boldsymbol D_{f}\boldsymbol \theta_{f}}_\text{revision error variance}+
  \underbrace{[\transp{\boldsymbol \delta}(\transp{\boldsymbol Z_{p}}\boldsymbol v-\transp{\boldsymbol Z}\boldsymbol \theta)]^{2}}_{bias^2}
}_\text{mean square revision error}+
\underbrace{2\transp{\boldsymbol l}(\transp{\boldsymbol U_{p}}\boldsymbol v-\transp{\boldsymbol U}\boldsymbol \theta)}_{\text{constraints}}.
}\end{equation} where \(\boldsymbol l\) is a vector of Lagrange
multipliers.

When \(\boldsymbol U=\boldsymbol X\) (and then
\(\boldsymbol Z = \boldsymbol 0\)), the constraint is equivalent to
preserving polynomials of degree \(d\) and we find asymmetric direct
filters (DAF) when \(\boldsymbol D=\sigma^2\boldsymbol K^{-1}\).

When \(\boldsymbol U=\transp{\begin{pmatrix}1&\cdots&1\end{pmatrix}}\),
\(\boldsymbol Z=\transp{\begin{pmatrix}-h&\cdots&+h\end{pmatrix}}\),
\(\boldsymbol \delta=\delta_1\), \(\boldsymbol D=\sigma^2\boldsymbol I\)
and when the symmetric filter is the Henderson filter, we find the
asymmetric Musgrave filters. This filter assumes that, for real-time
estimation, the data is generated by a linear process and that the
asymmetric filters preserve the constants
(\(\sum v_i=\sum \theta_i=1\)). These asymmetric filters depend on the
ratio \(\lvert\delta_1/\sigma\rvert\), which, assuming that the trend is
linear and the bias constant, can be linked to the I-C ratio
\(R=\frac{\bar{I}}{\bar{C}}=\frac{\sum\lvert I_t-I_{t-1}\rvert}{\sum\lvert C_t-C_{t-1}\rvert}\):
\(\delta_1/\sigma=2/(R\sqrt{\pi})\). This ratio is used in particular in
X-11 to determine the length of the Henderson filter. For monthly data:

\begin{itemize}
\item
  If the ratio is large (\(3.5< R\)) then a 23 term filter is used (to
  remove more noise).
\item
  If the ratio is small (\(R<1\)) a 9 term filter is used.
\item
  Otherwise (most of the cases) a 13-term filter is used.
\end{itemize}

When \(\boldsymbol U\) corresponds to the first \(d^*+1\) columns of
\(\boldsymbol X\), \(d^*<d\), the constraint is to reproduce polynomial
trends of degree \(d^*\). This introduces bias but reduces the variance.
Thus, \textcite{proietti2008} proposes three classes of asymmetric
filters:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \emph{Linear-Constant} (LC): \(y_t\) linear (\(d=1\)) and
  \(\boldsymbol v\) preserves constant signals (\(d^*=0\)). We obtain
  Musgrave filters when the Henderson kernel is used and the Henderson
  filter is used as the symmetric filter.
\item
  \emph{Quadratic-Linear} (QL): \(y_t\) quadratic (\(d=2\)) and
  \(\boldsymbol v\) preserves linear signals (\(d^*=1\)).
\item
  \emph{Cubic-Quadratic} (CQ): \(y_t\) cubic (\(d=3\)) and
  \(\boldsymbol v\) preserves quadratic signals (\(d^*=2\)).
\end{enumerate}

Supplemental materials shows the coefficients, gain and phase shift
functions of the four asymmetric filters.

\subsection{Extension with the timeliness
criterion}\label{sec-lptimeliness}

One drawback of the approach of \textcite{proietti2008} is the lack of
control over the phase shift. However, it is possible to improve the
modelling by incorporating in equation~\ref{eq-lppasym} the timeliness
criterion defined by \textcite{ch15HBSA}. This was proposed by Jean
Palate, then coded in Java and integrated into \texttt{rjd3filters}.

To measure the phase shift between the input series and the filtered
one, equation~\ref{eq-lppasym} introduces a criterion, \(T_g\), called
\emph{timeliness}. When a moving average \(\boldsymbol v\) is applied,
the level input signal is also altered by the gain function. That's why
the timeliness criterion depends on the gain and phase shift functions
(\(\rho_{\boldsymbol v}\) and \(\varphi_{\boldsymbol v}\)), the link
between both functions being made by a penalty function \(f\). The
criterion measures phase shift for the frequencies in the interval
\([\omega_1, \omega_2]\): \[
T_g(\boldsymbol v)=\int_{\omega_{1}}^{\omega_{2}}f(\rho_{\boldsymbol v}(\omega),\varphi_{\boldsymbol v}(\omega))\ud\omega
\] As mentioned in section~\ref{sec-propMM}, in the case of trend-cycle
extraction we can set \(\omega_1=0\) and for monthly series
\(\omega_2=2\pi/12.\) For the penalty function, the authors suggest
\(f\colon(\rho,\varphi)\mapsto\rho^2\sin(\varphi)^2\) and show thant in
that case the criteria can be written in a quadratic form: \[
T_g(\boldsymbol v) = \transp{\boldsymbol v}\boldsymbol T\boldsymbol v
\] with
\(T\boldsymbol=\int_{\omega_{1}}^{\omega_{2}}\boldsymbol{Si}(\omega)\transp{\boldsymbol{Si}(\omega)}\ud\omega\)
and \(\boldsymbol{Si}(\omega)=(\sin(p\omega),\dots, \sin (f\omega))\) if
\(\boldsymbol v\) uses \(p\) points in the past and \(f\) points in the
future.

Using the same notation as in section~\ref{sec-lppasymf},
\(\boldsymbol\theta\) represents the symmetric filter and
\(\boldsymbol v\) the asymmetric filter. With
\(\boldsymbol\theta=\transp{\begin{pmatrix}\transp{\boldsymbol\theta_p}&\transp{\boldsymbol\theta_f}\end{pmatrix}}\)
and \(\boldsymbol\theta_p\) of the same length as \(\boldsymbol v\), and
\(\boldsymbol g=\boldsymbol v-\boldsymbol \theta_p\), the timeliness
criterion can be expressed as: \[
T_g(\boldsymbol v)=\transp{\boldsymbol v}\boldsymbol T\boldsymbol v=\transp{\boldsymbol g}\boldsymbol T\boldsymbol g+2\transp{\boldsymbol \theta_p}\boldsymbol T\boldsymbol g+\transp{\boldsymbol \theta_p}\boldsymbol T\boldsymbol \theta_p
\quad\text{ with }\boldsymbol T\text{ a symmetric matrix}.
\] Furthermore, the objective function \(\varphi\) of
equation~\ref{eq-lppasym} can be rewritten: \begin{align*}
\varphi(\boldsymbol v)&=\transp{(\boldsymbol v-\boldsymbol \theta_p)}\boldsymbol D_{p}(\boldsymbol v-\boldsymbol \theta_p)+
  \transp{\boldsymbol \theta_f}\boldsymbol D_{f}\boldsymbol \theta_f+
  [\transp{\boldsymbol \delta}(\transp{\boldsymbol Z_{p}}\boldsymbol v-\transp{\boldsymbol Z}\boldsymbol \theta)]^{2}+
2\transp{\boldsymbol l}(\transp{\boldsymbol U_{p}}\boldsymbol v-\transp{\boldsymbol U}\boldsymbol \theta)\\
&=\transp{\boldsymbol g}\boldsymbol Q\boldsymbol g-2\boldsymbol P\boldsymbol g+2\transp{\boldsymbol l}(\transp{\boldsymbol U_{p}}\boldsymbol v-\transp{\boldsymbol U}\boldsymbol \theta)+\boldsymbol c,
\end{align*} with
\(\boldsymbol Q=\boldsymbol D_p+\boldsymbol Z_p\boldsymbol \delta\transp{\boldsymbol \delta}\transp{\boldsymbol Z}_p,\)
\(\boldsymbol P=\boldsymbol \theta_f\boldsymbol Z_f\boldsymbol \delta\transp{\boldsymbol \delta}\transp{\boldsymbol Z_p}\)
and \(\boldsymbol c\) a constant independent of \(\boldsymbol v.\)

Adding the timeliness criterion: \[
\widetilde\varphi(\boldsymbol v)=\transp{\boldsymbol g}\widetilde {\boldsymbol Q}\boldsymbol g-
2\widetilde {\boldsymbol P}\boldsymbol g+2\transp{\boldsymbol l}(\transp{\boldsymbol U_{p}}\boldsymbol v-\transp{\boldsymbol U}\boldsymbol \theta)+
\widetilde {\boldsymbol c},
\] where \(\alpha_T\) is the weight associated with the timeliness
criterion,
\(\widetilde {\boldsymbol Q}=\boldsymbol D_p+\boldsymbol Z_p\boldsymbol \delta\transp{\boldsymbol \delta}\transp{\boldsymbol Z_p} + \alpha_T\boldsymbol T\),
\(\widetilde {\boldsymbol P}=\boldsymbol \theta_f\boldsymbol Z_f\boldsymbol \delta\transp{\boldsymbol \delta}\transp{\boldsymbol Z_p}-\alpha_T\boldsymbol \theta_p\boldsymbol T\)
and \(\widetilde {\boldsymbol c}\) a constant independent of
\(\boldsymbol v.\)

With \(\alpha_T=0\) we find \(\varphi(\boldsymbol v)\). This extension
therefore makes it possible to find all the symmetric and asymmetric
filters presented in the previous section but also generalizes the
approach of \textcite{GrayThomson1996}. This extension is implemented in
the R function \texttt{rjd3filters::lp\_filter()}.

One drawback is that \(T_g(\boldsymbol v)\) is not normalized: the
weight \(\alpha_T\) has no economic sense. It is then difficult to
calibrate the value of \(\alpha_T\) since the value of
\(\lvert\delta/\sigma\rvert\) must also be defined. Hence, in this paper
we will not compare this extension to the other methods and we will
focus on the local calibration of \(\lvert\delta/\sigma\rvert\).
However, in future studies, we could imagine a two step calibration: fix
the value of \(\lvert\delta/\sigma\rvert\) to find \(\alpha_T\) by
cross-validation and then use this weight with a local parametrisation
of the ratio \(\lvert\delta/\sigma\rvert\) (as in
section~\ref{sec-localic}).

\subsection{Local parametrisation of asymmetric
filters}\label{sec-localic}

Asymmetric filters are usually parametrised globally:
\(\lvert\delta/\sigma\rvert\) is fixed using all the data (with the
IC-ratio or a cross-validation criterion). However, we might prefer a
local parametrisation: ratio \(\lvert\delta/\sigma\rvert\) which varies
as a function of time. Indeed, although the overall parametrisation is
generally valid, assuming a constant value of the
\(\lvert\delta/\sigma\rvert\) ratio for all asymmetric filters does not
appear relevant for real-time estimation, especially during periods of
economic downturn. For example, with the LC method, global
parametrisation means assuming that the slope of the trend is constant,
whereas during economic downturns it tends towards 0 up around the
turning point.

This is what is proposed in this article, with a local parametrisation
of the asymmetric filters by estimating \(\delta\) and \(\sigma^2\)
separately. Although this does not give an unbiased estimator of the
ratio \(\lvert\delta/\sigma\rvert\), it does allow the main evolutions
to be captured, such as the decay towards 0 before a turning point and
the growth after the turning point for the LC method:

\begin{itemize}
\item
  The variance \(\sigma^2\) can be estimated using the observed data set
  and the symmetric filter \((\theta_{-h},\dots,\theta_h)\). Indeed, as
  for example shown by \textcite{Loader1999}, the variance \(\sigma^2\)
  can be estimating using the normalized residual sum of squares: \[
  \hat\sigma^2=\frac{1}{(n-2h)-2\nu_1+\nu_2}\sum_{t=h+1}^{n-h}(y_t-\widehat{TC}_t)^2.
  \] Since the symmetric filter is used, only \(n-2h\) observations can
  be used to estimate \(\sigma^2\). \(\nu_1\) and \(\nu_2\) are two
  definitions of degrees of freedom of a local fit (generalization of
  the number of parameters of a parametric model). In the case of local
  polynomial smoothings, they can be linked to the coefficient of the
  associated filter: \(\nu_1 = (n-h)\theta_0\) (number of observations
  multiplied by the central coefficient of the moving average) and
  \(\nu_2 = (n-h)\sum_{i=-h}^{h} \theta_i^2\) (number of observations
  multiplied by the sum of squares of coefficients of the moving
  average). Thus, the variance can be estimated by: \[
  \hat\sigma^2=\frac{1}{n-2h}\sum_{t=h+1}^{n-h}\frac{(y_t-\widehat{TC}_t)^2}{1-2\theta_0^2+\sum_{i=-h}^{h} \theta_i^2}.
  \] This is implemented in the function
  \texttt{rjd3filters::var\_estimator()}. This estimator assumes that
  \(\widehat{TC}_t\) is unbiased, so variance estimates intervals are
  usually computed at small bandwidths (\(h\)). In this paper we study
  monthly series with small bandwidth (\(h=6\)), and the variance is
  estimated using local polynomial of degree 3 (Henderson filter), the
  biased of \(\widehat{TC}_t\) is thus already small and using smaller
  bandwidths have almost no impact on the estimates (see Supplemental
  materials).
\item
  The parameter \(\delta\) can be estimated by moving average from
  equation~\ref{eq-mmsym}. For example, for the LC method we can use the
  moving average
  \(\boldsymbol \theta_2=\boldsymbol K\boldsymbol X(\transp{\boldsymbol X}\boldsymbol K\boldsymbol X)^{-1}\boldsymbol e_{2}\)
  to obtain a local estimate of the slope and for the QL method we can
  use
  \(\boldsymbol \theta_3=\boldsymbol K\boldsymbol X(\transp{\boldsymbol X}\boldsymbol K\boldsymbol X)^{-1}\boldsymbol e_{3}\)
  to obtain a local estimate of the concavity. The DAF method then
  simply allows us to calculate the associated asymmetric moving
  averages.\\
  Although a moving average of different length to that used for
  estimating the trend could be considered, this seems to degrade the
  results in terms of phase shift (using the same methodology as in
  section~\ref{sec-comparison}). Furthermore, for the construction of
  moving averages, the trend can be modelled as being locally of degree
  2 or 3 (this has no impact on the final estimate of concavity). In
  this paper, we have chosen to model a trend of degree 2: this slightly
  reduces the phase shift (see Supplemental material) but slightly
  increases the revisions linked to the first estimate of the
  trend-cycle. figure~\ref{fig-mmpenteconcac} shows the moving averages
  used.
\end{itemize}

\begin{figure}[H]

\caption{\label{fig-mmpenteconcac}Moving averages used for real-time
estimation of slope and concavity.}

\centering{

\includegraphics[width=0.9\textwidth,height=\textheight]{img/filters_used/mm_penteconcavite.pdf}

}

\end{figure}%

To distinguish the gain from the local parametrisation of the noise
biased associated to the real-time estimates of the slope and concavity,
in the empirical applications of section~\ref{sec-comparison} we also
compares the results to the final local parametrisation obtained by
estimating \(\delta\) using all the data (i.e., using the symmetric
filters shown in figure~\ref{fig-mmpenteconcac}), while maintaining a
real-time estimate of \(\sigma^2\). figure~\ref{fig-mmpenteconcac-ex}
shows an example of the comparison of estimates of the ratio
\(|\delta_1/\sigma|\) with a global estimator (IC ratio) and two local
parametrisations:

\begin{itemize}
\item
  figure~\ref{fig-mmpenteconcac-ex-1}: using a symmetric filter to
  estimate \(\delta_1\) (final local parametrisation).
\item
  figure~\ref{fig-mmpenteconcac-ex-2}: using the asymmetric filter which
  need two points in the future to estimate \(\delta_1\). This is closed
  to the real-time estimates since two points in the future are needed
  to detect a turning point.
\end{itemize}

The turning points are clearly detected by the local estimators
(changement in the slope), when \(|\delta_1/\sigma|\) is estimated in
the real-time estimate. However, the real-time local estimates of
\(|\delta_1/\sigma|\) are noisy: the varability of this estimates, used
to build asymmetric filters, will then lead to another source of
revisions in the intermediate estimates.

\begin{figure}[H]

\caption{\label{fig-mmpenteconcac-ex}Comparison of the estimates of
\(|\delta_1/\sigma|\) with a local estimator and global estimator (with
IC ratio) on a simulated series, the vertical lines being the simulated
turning points.}

\begin{minipage}{\linewidth}

\subcaption{\label{fig-mmpenteconcac-ex-1}Symmetric moving average used
for \(\hat \delta_1\).}

\centering{

\includegraphics[width=0.9\textwidth,height=\textheight]{img/filters_used/mm_penteconcavite_ex.pdf}

}

\end{minipage}%
\newline
\begin{minipage}{\linewidth}

\subcaption{\label{fig-mmpenteconcac-ex-2}Asymmetric moving average
using two future points (\(q=2\)) used for \(\hat \delta_1\).}

\centering{

\includegraphics[width=0.9\textwidth,height=\textheight]{img/filters_used/mm_penteconcavite_ex2.pdf}

}

\end{minipage}%

\end{figure}%

There is no function in \texttt{rjd3filters} to directly compute those
moving averages but they can be easily computed using matrix computation
formula and applied to the series rjd3filters`.

\section{Comparison of different methods}\label{sec-comparison}

The different methods are compared on simulated and real data. For all
series, a symmetric 13-term filter is used: we then assume than the
bandwidth is fixed, as in the seasonal adjustement method X-11. A
nearest neighbour bandwidth is also tested (13-term asymmetric filters:
real-time estimates uses observation between \(y_{t-12}\) and \(y_{t}\),
the second asymmetric filters uses obversations between \(y_{t-11}\) and
\(y_{t+1}\), etc.) but with no improvement in the results (see
supplemental material). More complexe methods to locally select
bandwidth, like the one proposed by \textcite{fan1992variable}, were not
tested. All the methods to create asymmetric filters are also compared
with estimates obtained by extending the series with an ARIMA model,
then applying a 13-term symmetric Henderson filter. The ARIMA model is
determined automatically, using \textcite{autoarima} algorithm (function
\texttt{forecast::auto.arima()}) on the last 12 years, with no seasonal
lag (the series being seasonally adjusted) and no external variables
(such as external regressors for correction of atypical points).

The performance of the different methods is judged on criteria relating
to revisions (between two consecutive estimates and in relation to the
final estimate) and the number of periods required to detect turning
points.

\subsection{Simulated series}\label{simulated-series}

\subsubsection{Methodology}\label{methodology}

Following a methodology close to that of \textcite{DarneDagum2009}, nine
monthly series are simulated between January 1960 and December 2020 with
different levels of variability. Each simulated series
\(y_t= C_t+ T_t + I_t\) can be written as the sum of three components:

\begin{itemize}
\item
  the cycle
  \(C_t = \rho [\cos (2 \pi t / \lambda) +\sin (2 \pi t / \lambda)]\),
  \(\lambda\) is fixed at 72 (6-year cycles, so there are 19 detectable
  turning points);
\item
  the trend \(T_t = T_{t-1} + \nu_t\) with
  \(\nu_t \sim \mathcal{N}(0, \sigma_\nu^2)\), \(\sigma_\nu\) being
  fixed at \(0.08\);
\item
  and the irregular \(I_t = e_t\) with
  \(e_t \sim \mathcal{N}(0, \sigma_e^2)\).
\end{itemize}

For the different simulations, we vary the parameters \(\rho\) and
\(\sigma_e^2\) in order to obtain series with different signal-to-noise
ratios (see figure~\ref{fig-graphs-data-simul}):

\begin{itemize}
\item
  High signal-to-noise ratio (i.e.~low I-C ratio and low variability):
  \(\sigma_e^2=0.2\) and \(\rho = 3.0, 3.5\) or \(4.0\) (I-C ratio
  between 0.9 and 0.7);
\item
  Medium signal-to-noise ratio (i.e.~medium I-C ratio and medium
  variability): \(\sigma_e^2=0.3\) and \(\rho = 1.5,\, 2.0\) or \(3.0\)
  (I-C ratio between 2.3 and 1.4);
\item
  Low signal-to-noise ratio (i.e.~high I-C ratio and high variability):
  \(\sigma_e^2=0.4\) and \(\rho = 0.5,\, 0.7\) or \(1.0\) (I-C ratio
  between 8.9 and 5.2).
\end{itemize}

\begin{figure}[H]

\caption{\label{fig-graphs-data-simul}Simulated series with low
(\(\sigma_e^2=0.2\) and \(\rho = 3.55\)), medium (\(\sigma_e^2=0.3\) and
\(\rho = 2.0\)) and high variability (\(\sigma_e^2=0.4\) and
\(\rho = 1.0\)).}

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{img/simulations/simul_data.pdf}

}

\end{figure}%

For each series and each date, the trend-cycle is estimated using the
different methods presented in this report. Three quality criteria are
also calculated:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Computation of the phase shift in the detection of turning points. In
  this paper, we focus on turning points associated to business cycle:
  it is defined as the succession of phases of economic recession and
  expansion, delimited by peaks (highest level of activity) and troughs
  (lowest level of activity), see for example \textcite{ferrara} for a
  description of the different economic cycles. In this paper, following
  \textcite{Zellner1991}, we used an accepted definition for
  turning-points:

  \begin{itemize}
  \tightlist
  \item
    An upturn occurs when the economy moves from a phase of recession to
    a phase of expansion. This is the case at date \(t\) when
    \(TC_{t-3}\geq TC_{t-2}\geq TC_{t-1}<TC_t\leq TC_{t+1}\).\\
  \item
    A downturn occurs when the economy moves from a phase of expansion
    to a phase of recession. This is the case at date \(t\) when
    \(TC_{t-3}\leq TC_{t-2}\leq TC_{t-1}>TC_t\geq TC_{t+1}\).
  \end{itemize}

  Lets denote \(TC_{t|t'}\) the estimate of the trend-cycle at date
  \(t\) using the data up to date \(t'\). The phase shift is often
  defined as the number of months required to detect the right turning
  point: it would then be equal to two if, in the case of a downturn,
  \(TC_{t-3|t+1}\leq TC_{t-2|t+1}\leq TC_{t-1|t+1}>TC_{t|t+1}\geq TC_{t+1|t+1}\)
  (two observations, at dates \(t\) and \(t+1\), are needed to qualify
  the date \(t\) as turning point). Note that, in that case, the phase
  shift could be equivalently set to 1 (only one observation after \(t\)
  is needed to define \(t\) as a turning point), without any difference
  on the comparison of the different methods.\\
  In this paper we use a slightly modified criterion: the phase shift is
  defined as the number of months needed to detect the right turning
  point without any future revision. Thus, in the case of a downturn,
  the phase shift is two if
  \(TC_{t-3|t+1}\leq TC_{t-2|t+1}\leq TC_{t-1|t+1}>TC_{t|t+1}\geq TC_{t+1|t+1}\)
  and, for all dates \(t'\) higher than \(t+1\):
  \(TC_{t-3|t'}\leq TC_{t-2|t'}\leq TC_{t-1|t'}>TC_{t|t'}\geq TC_{t+1|t'}\).
  Since a 13-term Henderson filter is used for the final estimates of
  the trend-cycle, the phase shift is at most height (the final estimate
  of \(TC_{t+1}\) is compute a the date \(t+7\)). This definition is
  used because it can happen that the right turning point is detected by
  asymmetric filters but is not detected by the final estimate using a
  symmetric filter (this is the case for 41 reversal points out of the 9
  series with asymmetric Musgrave filters) or that there are revisions
  in successive estimates (this is the case for 7 turning points out of
  the 9 series with asymmetric Musgrave filters). Finally, relatively
  few turning points are detected at the right date with the final
  estimate. With the 13-term Henderson filter, 18 are correctly detected
  in series with low variability (out of 57 possible), 11 in series with
  medium variability and 12 in series with high variability.
\item
  The average of the relative deviations between to the last estimate
  (comparison of the \(q\)\textsuperscript{th} estimate and the last
  estimate): \[
  MAE_{fe}(q)=\mathbb E\left[
  \left|\frac{
  TC_{t|t+q} -  TC_{t|last}
  }{
   TC_{t|last}
  }\right|
  \right].
  \]
\item
  The average of the relative deviations between two consecutive
  estimates (the \(q\)\textsuperscript{th} estimate and the
  \(q+1\)\textsuperscript{th} estimate): \[
  MAE_{ce}(q)=\mathbb E\left[
  \left|\frac{
  TC_{t|t+q} - TC_{t|t+q+1}
  }{
  TC_{t|t+q+1}
  }\right|
  \right].
  \]
\end{enumerate}

The simulated series are of length 60 years, which is often not
realistic for economic time series. However, since all the methods are
local, the results are not affected by the length of the series. The
length of the series would only have an impact for the identification
and the estimation of the ARIMA model. In this case, since the same data
generating process is used during the 60 years, it could be relevant to
identify the ARIMA model using all the data. Even if it would improve
the results in term of phase shift (see supplemental material) we prefer
to only use the last 12 years to identify and estimate the ARIMA model,
to be closer to what would have been done with a real time series.

\subsection{Comparison}\label{comparison}

Excluding for the moment the local parametrisations of the polynomial
filters, the linear-constant polynomial filter (LC) seems to give the
best results in terms of delay in the detection of turning points
(figure~\ref{fig-graphstpsimul}). Performance is relatively close to
that obtained by extending the series using an ARIMA model. However,
when variability is low, the LC filter seems to give poorer results and
the quadratic-linear polynomial filter (QL) seems to give the best
results.

For series with moderate variability, local parametrisation of the LC
and QL filters reduces the phase shift. For series with high
variability, the phase shift is only reduced by using the final
\(\hat\delta\) parameters: real-time estimates seem to add more
variance. For series with low variability, performance seems to be
slightly improved only with the LC filter.

\begin{figure}

\caption{\label{fig-graphstpsimul}Distribution of phase shift on
simulated series.}

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{img/simulations/phase_shift_simul.pdf}

Lecture: this plots represent box plots of the phase shift: the box
represents the interquartile range, the central horizontal line the
median the dot the outliers and the whisker the range of the data (if no
outlier).\\
For simulated series with medium variability (when a symmetric filter of
length 13 is relevant), with the linear constant (LC) method, 50\% of
the turning point are detected with a phase shift of at most 5 months
whereas when the LC filter is parametrised locally (using), 50\% of the
turning point are detected with a phase shift of at most 4 months. For
the direct asymmetric filter (DAF), 25\% of the turning are detected
with a phase shift of at least 7 months (compares to 5 for the LC
method).

}

\end{figure}%

In terms of revisions, the variability of the series has little effect
on the respective performances of the different methods, but it does
affect the orders of magnitude, which is why the results are presented
only for series with medium variability (table~\ref{tbl-simulrev}). In
general, LC filters always minimise revisions (with relatively small
effect from the local parametrisation of the filters) and revisions are
greater with cubic-quadratic (CQ) and direct (DAF) polynomial filters.

For the QL filter, there is a large revision between the second and
third estimates: this may be due to the fact that for the second
estimate (when one point in the future is known), the QL filter assigns
a greater weight to the estimate in \(t+1\) than to the estimate in
\(t\), which creates a discontinuity. This revision is considerably
reduced by parametrising the filter locally. For polynomial filters
other than the LC filter, the large revisions at the first estimate were
to be expected given the coefficient curve: a very large weight is
associated with the current observation and there is a strong
discontinuity between the moving average used for real-time estimation
(when no point in the future is known) and the other moving averages.

Extending the series using an ARIMA model gives revisions with the
latest estimates of the same order of magnitude as the LC filter, but
slightly larger revisions between consecutive estimates, particularly
between the fourth and fifth estimates (as might be expected, as
highlighted in section~\ref{sec-mmasym}).

\begin{longtable}[t]{ccccccc}

\caption{\label{tbl-simulrev}Average of the relative deviations of the
revisions for the different filters on simulated series with medium
variability.}

\tabularnewline

\toprule
Method & \$q=0\$ & \$q=1\$ & \$q=2\$ & \$q=3\$ & \$q=4\$ & \$q=5\$\\
\midrule
\addlinespace[0.3em]
\multicolumn{7}{l}{\textbf{$MAE_{fe}(q) = \mathbb E\left[\left|(TC_{t|t+q} -  TC_{t|last})/TC_{t|last}\right|\right]$}}\\
\hspace{1em}LC & 0.21 & 0.10 & 0.03 & 0.03 & 0.03 & 0.01\\
\hspace{1em}LC local param. (final estimates) & 0.19 & 0.09 & 0.03 & 0.03 & 0.03 & 0.01\\
\hspace{1em}LC local param. & 0.29 & 0.10 & 0.03 & 0.03 & 0.03 & 0.01\\
\hspace{1em}QL & 0.33 & 0.10 & 0.04 & 0.04 & 0.03 & 0.01\\
\hspace{1em}QL local param. (final estimates) & 0.21 & 0.10 & 0.03 & 0.03 & 0.03 & 0.01\\
\hspace{1em}QL local param. & 0.30 & 0.10 & 0.04 & 0.03 & 0.03 & 0.01\\
\hspace{1em}CQ & 0.45 & 0.13 & 0.13 & 0.09 & 0.06 & 0.02\\
\hspace{1em}DAF & 0.47 & 0.15 & 0.15 & 0.09 & 0.06 & 0.02\\
\hspace{1em}ARIMA & 0.22 & 0.10 & 0.03 & 0.03 & 0.03 & 0.01\\
\addlinespace[0.3em]
\multicolumn{7}{l}{\textbf{$MAE_{ce}(q)=\mathbb E\left[
\left|(TC_{t|t+q} - TC_{t|t+q+1})/TC_{t|t+q+1}\right|
\right]$}}\\
\hspace{1em}LC & 0.19 & 0.10 & 0.02 & 0.01 & 0.07 & 0.01\\
\hspace{1em}LC local param. (final estimates) & 0.20 & 0.10 & 0.03 & 0.01 & 0.05 & 0.01\\
\hspace{1em}LC local param. & 0.24 & 0.11 & 0.03 & 0.01 & 0.05 & 0.01\\
\hspace{1em}QL & 0.29 & 3.46 & 0.00 & 0.03 & 0.04 & 0.01\\
\hspace{1em}QL local param. (final estimates) & 0.31 & 0.11 & 0.02 & 0.01 & 0.04 & 0.01\\
\hspace{1em}QL local param. & 0.24 & 0.16 & 0.00 & 0.03 & 0.04 & 0.01\\
\hspace{1em}CQ & 0.43 & 0.02 & 0.10 & 0.07 & 0.05 & 0.02\\
\hspace{1em}DAF & 0.66 & 0.24 & 0.11 & 0.14 & 0.06 & 0.02\\
\hspace{1em}ARIMA & 0.23 & 0.10 & 0.02 & 0.03 & 0.25 & 0.01\\
\bottomrule

\end{longtable}

Lecture: the estimates for \(q=0\) correspond to the first estimates of
the trend-cycle (real-time estimates when 0 point in the future is
known), \(q=1\) correspond to the second estimates of the trend-cycle
(when 1 point in the future is known). Since the final estimates are
obtain using a symmetric filter of bandwidth 6, \(q=6\) corresponds to
the final estimates of the trend-cycle (when 6 points in the future are
known). \(MAE_{fe}(0)\) compares the revision between the first estimate
(\(q=0\)) and the last estimate (\(q=6\)) and \(MAE_{ce}(0)\) compares
the revision between the first estimate (\(q=0\)) and the second
estimate (\(q=1\)).

\subsection{Real series}\label{real-series}

The differences between the methods are also illustrated using an
example from the FRED-MD database (\textcite{fredmd}) containing
economic series on the United States. The series studied correspond to
the database published in November 2022. The series studied is the level
of employment in the United States (series \texttt{CE160V}, used in
logarithm) around the February 2001 turning point, consistent with the
monthly dating of the turning points. This point and this series were
chosen because the economic turnaround is particularly visible in the
raw series (figure~\ref{fig-ce16ov-previmp-lp}), because this series is
used for dating economic cycles and because the FRED-MD database
facilitates the reproducibility of the results (thanks to the
availability of series published on past dates). Studying the series up
to January 2020, this series has a medium variability: a symmetric
filter of 13 terms is therefore appropriate.

figure~\ref{fig-ce16ovlp} shows the successive estimates of the
trend-cycle using the different methods studied.

For this series, the phase shift is 6 months for the LC and CQ methods
and the extension of the series by ARIMA. Except for the LC method
(where the turning point is first detected in January 2001), the high
phase shift is due to revisions in the turning point detected in the
intermediate estimates (even if the correct turning point is detected
after 2 months, it is no longer detected after 3 months). It is of two
months for the other methods (QL, DAF).

In this example, local parametrisation does not reduce the phase shift,
but it does reduce revisions. The QL and DAF polynomials lead to greater
variability in the intermediate estimates, especially in February 2001.

\begin{figure}[H]

\caption{\label{fig-ce16ovlp}Successive estimates of the trend-cycle in
US employment (in logarithms).}

\centering{

\includegraphics[width=0.95\textwidth,height=\textheight]{img/nber/ce16ov_fev2001_lp.pdf}

}

\end{figure}%

The quality of the intermediate estimates can also be analysed using the
implicit forecasts of the different methods
(figure~\ref{fig-ce16ov-previmp-lp}). As a reminder, these are forecasts
of the raw series which, by applying the symmetric Henderson filter to
the extended series, give the same estimates as the asymmetric moving
averages. The forecasts of the ARIMA model are naive and do not take the
turning point into account, unlike the other methods. Finally, local
parametrisation of the QL filter produces much more consistent
forecasts.

\begin{figure}[H]

\caption{\label{fig-ce16ov-previmp-lp}Implicit forecasts linked to
successive estimates of the trend-cycle in US employment (in logarithms)
using local polynomial methods.}

\centering{

\includegraphics[width=0.95\textwidth,height=\textheight]{img/nber/ce16ov_fev2001_prev_imp_lp.pdf}

}

\end{figure}%

\section{Discussion}\label{discussion}

\subsection{Outliers and COVID-19}\label{outliers-and-covid-19}

In this article and in those associated to trend-cycle extraction
techniques, the asymmetrical filters are applied and compared on series
already seasonally adjusted or without seasonality (simulated series).
The revisions and therefore the phase shift are limited to 6 months
(when the symmetrical filter is 13 terms), and this has the advantage of
isolating the impacts of the different filters from the other processes
inherent in seasonal adjustment. There are, however, two drawbacks to
this simplification:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Firstly, the estimation of the seasonally-adjusted series depends on
  the method used to extract the trend-cycle. The choice of method used
  to estimate the trend-cycle can therefore have an impact well beyond 6
  months.
\item
  On the other hand, as moving averages are linear operators, they are
  sensitive to the presence of atypical points. Direct application of
  the methods can therefore lead to biased estimates, due to their
  presence, whereas seasonal adjustment methods (such as the
  X-13ARIMA-SEATS method) have a correction module for atypical points.
  Furthermore, as shown in particular by \textcite{dagum1996new}, the
  final symmetrical filter used by X-13ARIMA to extract the trend-cycle
  (and therefore the one indirectly used when applying the methods to
  seasonally-adjusted series) misses around 72\% of 9- or 10-month
  cycles (generally associated with noise rather than the trend-cycle).
  The final asymmetrical filters even amplify 9 or 10-month cycles. This
  can result in the introduction of undesirable ripples, i.e.~the
  detection of false turning points. This problem is reduced by
  correcting atypical points (these cycles being considered irregular).
  This is how the \emph{Nonlinear Dagum Filter} (NLDF) was developed:

  \begin{enumerate}
  \def\labelenumii{\alph{enumii}.}
  \item
    applying the X-13ARIMA atypical point correction algorithm to the
    seasonally adjusted series, then extending it with an ARIMA model;
  \item
    perform a new atypical point correction using a much stricter
    threshold and then apply the 13-term symmetrical filter. Assuming a
    normal distribution, this amounts to modifying 48\% of the irregular
    values.
  \end{enumerate}

  The \emph{cascade linear filter} (CLF), studied in particular in
  \textcite{dagumBianconcini2023}, corresponds to an approximation of
  the NLDF using a 13-term filter and when the forecasts are obtained
  from an ARIMA(0,1,1) model where \(\theta=0.40.\).
\end{enumerate}

One possible direction of study would then be to investigate more
precisely the effect of atypical points on trend-cycle estimation and
the detection of turning points, but also to explore new types of
asymmetric filters based on robust methods (such as robust local
regressions, moving medians, etc.).

\begin{figure}[H]

\caption{\label{fig-ce16ovcovid}Successive estimates of the trend-cycle
in US employment (in logarithms) during the COVID-19.}

\centering{

\includegraphics[width=0.95\textwidth,height=\textheight]{img/nber/ce16ov_covid_lp.pdf}

}

\end{figure}%

\begin{figure}[H]

\caption{\label{fig-ce16ov-previmp-covid}Implicit forecasts linked to
successive estimates of the trend-cycle in US employment (in logarithms)
using local polynomial methods.}

\centering{

\includegraphics[width=0.95\textwidth,height=\textheight]{img/nber/ce16ov_covid_prev_imp_lp.pdf}

}

\end{figure}%

\subsection{Other frequencies}\label{other-frequencies}

\section{Conclusion}\label{conclusion}

For business cycle analysis, most statisticians use trend-cycle
extraction methods, either directly or indirectly. They are used, for
example, to reduce the noise of an indicator in order to improve its
analysis, and models, such as forecasting models, usually use seasonally
adjusted series based on these methods.

This paper presents the R package \texttt{rjd3filters}, which implements
several methods to build moving averages for real-time trend-cycle
estimation. It also offers various functions for the theoretical
analysis of moving averages (gain functions, phase shift, quality
criteria, etc.) and judging the quality of the latest estimates (for
example with implicit forecasts). It also makes it easy to combine
moving averages and also to integrate custom moving averages into the
X-11 seasonal adjustment algorithm (with the
\texttt{rjd3filters::x11plus\_trend()} function). \texttt{rjd3filters}
thus facilitates research into the use of moving averages as part of
real-time cycle trend estimation. This is the case, for example, with
the local parametrisation of Musgrave moving averages and other
asymmetric moving averages linked to local polynomial regression,
presented in this article.

By comparing the different methods, we can learn a few lessons about the
construction of these moving averages.

During economic downturns, asymmetric filters used as an alternative to
extending the series using the ARIMA model can reduce revisions to
intermediate estimates of the trend-cycle and enable turning points to
be detected more quickly.

At the end of the period, modelling polynomial trends of degree greater
than three (cubic-quadratic, CQ, and direct, DAF) seems to introduce
variance into the estimates (and therefore more revisions) without
allowing faster detection of turning points. At the end of the period,
for estimating the trend-cycle, we can therefore restrict ourselves to
methods modelling polynomial trends of degree two or less
(linear-constant, LC, and quadratic-linear, QL). In addition,
parametrising polynomial filters locally as suggested in this paper
enables turning points to be detected more quickly (especially for the
QL filter). Even when the phase shift is not reduced, local
parametrisation is recommended because it reduces revisions and produces
intermediate estimates that are more consistent with expected future
trends. However, with these methods, the length of the filter used must
be adapted to the variability of the series: if the filter used is too
short (i.e.~if the variability of the series is ``low''), retaining
polynomial trends of degree one or less (LC method) produces poorer
results in terms of detecting turning points.

This study could be extended in many ways. One possible extension would
be to look at the impact of filter length on the detection of turning
points. Asymmetric filters are calibrated using indicators calculated
for the estimation of symmetric filters (for example, to automatically
determine their length), whereas a local estimate might be preferable.
Furthermore, we have only focused on monthly series with a 13-term
symmetric filter, but the results may be different if the symmetric
filter studied is longer or shorter and if we study series with other
frequencies (quarterly or daily, for example).

Another possibility would be to study the impact of atypical points:
moving averages, like any linear operator, are very sensitive to the
presence of atypical points. To limit their impact, in X-11 a strong
correction for atypical points is performed on the irregular component
before applying the filters to extract the trend-cycle. This leads us to
study the impact of these points on the estimation of the trend-cycle
and turning points, and also to explore new types of asymmetric filters
based on robust methods (such as robust local regressions or moving
medians).

\newpage

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\printbibliography[heading=none]

\end{document}