### 2024-04-25 paper/supp_material.qmd

---
title: "Supplemental material"
format:
  pdf:
    documentclass: article
    number-sections: true
    pdf-engine: pdflatex
    toc: false
    include-in-header: [preamble.tex]
    #include-before-body: template/preambule.tex
    geometry:
      - top=2cm
      - bottom=2cm
      - left = 2.5cm
      - right = 2.5cm
linestretch: 2
echo: false
tbl-cap-location: top
fig-cap-location: top
---

\appendix

# Coefficients, gain and phase shift functions {#sec-an-cof}

```{r, out.width="100%"}
#| label: fig-graphs-coef-lc
#| fig-cap: Coefficients, gain and phase shift functions for the Linear-Constant (LC) filter with $I/C=3.5$.
img <- sprintf("img/filters_used/%s.%s", tolower("LC"), "pdf")
knitr::include_graphics(img)
```

```{r, out.width="100%"}
#| label: fig-graphs-coef-ql
#| fig-cap: Coefficients, gain and phase shift functions for the Quadratic-Linear (QL) filter with $I/C=3.5$.
img <- sprintf("img/filters_used/%s.%s", tolower("QL"), "pdf")
knitr::include_graphics(img)
```

```{r, out.width="100%"}
#| label: fig-graphs-coef-cq
#| fig-cap: Coefficients, gain and phase shift functions for the Cubic-Quadratic (CQ) filter with $I/C=3.5$.
img <- sprintf("img/filters_used/%s.%s", tolower("CQ"), "pdf")
knitr::include_graphics(img)
```

```{r, out.width="100%"}
#| label: fig-graphs-coef-daf
#| fig-cap: Coefficients, gain and phase shift functions for the direct asymmetric filter (DAF).
img <- sprintf("img/filters_used/%s.%s", tolower("DAF"), "pdf")
knitr::include_graphics(img)
```

\newpage

# Different specifications for the estimation of the slope and concavity

::: {#fig-graphs-lc-deg-final}
```{r, echo=FALSE, out.width="100%"}
knitr::include_graphics("img/simulations/phase_shift_simul_local_final_lc.pdf")
```

Distribution of phase shift associated to the local parametrisation of the linear-constant (LC) filter using the final estimates of $\delta/\sigma$, by the bandwidth $h$ of the henderson filter used to estimate $\sigma$ and the degree $d$ of the trend to estimate $\delta$.
:::

::: {#fig-graphs-lc-deg}
```{r, echo=FALSE, out.width="100%"}
knitr::include_graphics("img/simulations/phase_shift_simul_local_lc.pdf")
```

Distribution of phase shift associated to the local parametrisation of the linear-constant (LC) filter using the final estimates of $\delta/\sigma$, by the bandwidth $h$ of the henderson filter used to estimate $\sigma$ and the degree $d$ of the trend to estimate $\delta$.
:::

::: {#fig-graphs-ql-deg-final}
```{r, echo=FALSE, out.width="100%"}
knitr::include_graphics("img/simulations/phase_shift_simul_local_final_ql.pdf")
```

Distribution of phase shift associated to the local parametrisation of the quadratic-linear (QL) filter using real time estimates of $\delta/\sigma$, by the bandwidth $h$ of the henderson filter used to estimate $\sigma$ and the degree $d$ of the trend to estimate $\delta$.
:::

::: {#fig-graphs-ql-deg}
```{r, echo=FALSE, out.width="100%"}
knitr::include_graphics("img/simulations/phase_shift_simul_local_ql.pdf")
```

Distribution of phase shift associated to the local parametrisation of the quadratic-linear (QL) filter using real time estimates of $\delta/\sigma$, by the bandwidth $h$ of the henderson filter used to estimate $\sigma$ and the degree $d$ of the trend to estimate $\delta$.
:::

\newpage

# Impact of the span used to detect the ARIMA model

::: {#fig-graphs-impact-span}
```{r, echo=FALSE, out.width="100%"}
knitr::include_graphics("img/simulations/phase_shift_simul_arima_length.pdf")
```

Distribution of phase shift on simulated series by number of years used to identify and estimate the ARIMA model.
:::

\newpage
# Nearest neighbour bandwidth

```{r, out.width="100%"}
#| label: fig-graphs-coef-lc-nn
#| fig-cap: Coefficients, gain and phase shift functions for the Linear-Constant (LC) filter with $I/C=3.5$ and a nearest neighbour bandwidth.
img <- sprintf("img/filters_used/%s_nn.%s", tolower("LC"), "pdf")
knitr::include_graphics(img)
```

```{r, out.width="100%"}
#| label: fig-graphs-coef-ql-nn
#| fig-cap: Coefficients, gain and phase shift functions for the Quadratic-Linear (QL) filter with $I/C=3.5$ and a nearest neighbour bandwidth.
img <- sprintf("img/filters_used/%s_nn.%s", tolower("QL"), "pdf")
knitr::include_graphics(img)
```

```{r, out.width="100%"}
#| label: fig-graphs-coef-cq-nn
#| fig-cap: Coefficients, gain and phase shift functions for the Cubic-Quadratic (CQ) filter with $I/C=3.5$ and a nearest neighbour bandwidth.
img <- sprintf("img/filters_used/%s_nn.%s", tolower("CQ"), "pdf")
knitr::include_graphics(img)
```

```{r, out.width="100%"}
#| label: fig-graphs-coef-daf-nn
#| fig-cap: Coefficients, gain and phase shift functions for the direct asymmetric filter (DAF) with a nearest neighbour bandwidth.
img <- sprintf("img/filters_used/%s_nn.%s", tolower("DAF"), "pdf")
knitr::include_graphics(img)
```

::: {#fig-nn-lp}
```{r, echo=FALSE, out.width="100%"}
knitr::include_graphics("img/simulations/phase_shift_simul_nn_lp.pdf")
```

Distribution of phase shift using nearest neighbour bandwidth.
:::

::: {#fig-nn-localparam}
```{r, echo=FALSE, out.width="100%"}
knitr::include_graphics("img/simulations/phase_shift_simul_nn_lp_localparam.pdf")
```
Note: nearest bandwidth moving average are used to estimate slope and concavity.
We thus do not distinguish local parametrisation and final local parametrisation which asumptions might not be plausible.
Indeed, using the same framework as for fixed bandwidth, the final local parametrisation for real time estimates would implie modelling a polynomial of degree 2 or 3 other 25 months (which doesn't seem plausible).

Distribution of phase shift using nearest neighbour bandwidth for polynomial filters with local parametrisation.
:::

\newpage

# Quarterly data

::: {#fig-graphs-quarterly}
```{r, echo=FALSE, out.width="100%"}
knitr::include_graphics("img/simulations/phase_shift_simul_q.pdf")
```

Distribution of phase shift on quarterly data.
:::