### 2024-04-29 paper/supp_material.qmd

---
title: "Supplemental material"
format:
  pdf:
    documentclass: article
    number-sections: true
    pdf-engine: pdflatex
    toc: false
    include-in-header: [preamble.tex]
    #include-before-body: template/preambule.tex
    geometry:
      - top=2cm
      - bottom=2cm
      - left = 2.5cm
      - right = 2.5cm
linestretch: 2
echo: false
tbl-cap-location: top
fig-cap-location: top
---

\appendix

```{r}
#| include: false
library(kableExtra)

```

# Coefficients, gain and phase shift functions {#sec-an-cof}

```{r, out.width="100%"}
#| label: fig-graphs-coef-lc
#| fig-cap: Coefficients, gain and phase shift functions for the Linear-Constant (LC) filter with $I/C=3.5$.
img <- sprintf("img/filters_used/%s.%s", tolower("LC"), "pdf")
knitr::include_graphics(img)
```

```{r, out.width="100%"}
#| label: fig-graphs-coef-ql
#| fig-cap: Coefficients, gain and phase shift functions for the Quadratic-Linear (QL) filter with $I/C=3.5$.
img <- sprintf("img/filters_used/%s.%s", tolower("QL"), "pdf")
knitr::include_graphics(img)
```

```{r, out.width="100%"}
#| label: fig-graphs-coef-cq
#| fig-cap: Coefficients, gain and phase shift functions for the Cubic-Quadratic (CQ) filter with $I/C=3.5$.
img <- sprintf("img/filters_used/%s.%s", tolower("CQ"), "pdf")
knitr::include_graphics(img)
```

```{r, out.width="100%"}
#| label: fig-graphs-coef-daf
#| fig-cap: Coefficients, gain and phase shift functions for the direct asymmetric filter (DAF).
img <- sprintf("img/filters_used/%s.%s", tolower("DAF"), "pdf")
knitr::include_graphics(img)
```

\newpage

# Different specifications for the estimation of the slope and concavity

::: {#fig-graphs-lc-deg-final}
```{r, echo=FALSE, out.width="100%"}
knitr::include_graphics("img/simulations/phase_shift_simul_local_final_lc.pdf")
```

Distribution of phase shift associated to the local parametrisation of the linear-constant (LC) filter using the final estimates of $\delta/\sigma$, by the bandwidth $h$ of the henderson filter used to estimate $\sigma$ and the degree $d$ of the trend to estimate $\delta$.
:::

::: {#fig-graphs-lc-deg}
```{r, echo=FALSE, out.width="100%"}
knitr::include_graphics("img/simulations/phase_shift_simul_local_lc.pdf")
```

Distribution of phase shift associated to the local parametrisation of the linear-constant (LC) filter using the final estimates of $\delta/\sigma$, by the bandwidth $h$ of the henderson filter used to estimate $\sigma$ and the degree $d$ of the trend to estimate $\delta$.
:::

::: {#fig-graphs-ql-deg-final}
```{r, echo=FALSE, out.width="100%"}
knitr::include_graphics("img/simulations/phase_shift_simul_local_final_ql.pdf")
```

Distribution of phase shift associated to the local parametrisation of the quadratic-linear (QL) filter using real time estimates of $\delta/\sigma$, by the bandwidth $h$ of the henderson filter used to estimate $\sigma$ and the degree $d$ of the trend to estimate $\delta$.
:::

::: {#fig-graphs-ql-deg}
```{r, echo=FALSE, out.width="100%"}
knitr::include_graphics("img/simulations/phase_shift_simul_local_ql.pdf")
```

Distribution of phase shift associated to the local parametrisation of the quadratic-linear (QL) filter using real time estimates of $\delta/\sigma$, by the bandwidth $h$ of the henderson filter used to estimate $\sigma$ and the degree $d$ of the trend to estimate $\delta$.
:::

\newpage

# Impact of the span used to detect the ARIMA model

::: {#fig-graphs-impact-span}
```{r, echo=FALSE, out.width="100%"}
knitr::include_graphics("img/simulations/phase_shift_simul_arima_length.pdf")
```

Distribution of phase shift on simulated series by number of years used to identify and estimate the ARIMA model.
:::

\newpage
# Nearest neighbour bandwidth

```{r, out.width="100%"}
#| label: fig-graphs-coef-lc-nn
#| fig-cap: Coefficients, gain and phase shift functions for the Linear-Constant (LC) filter with $I/C=3.5$ and a nearest neighbour bandwidth.
img <- sprintf("img/filters_used/%s_nn.%s", tolower("LC"), "pdf")
knitr::include_graphics(img)
```

```{r, out.width="100%"}
#| label: fig-graphs-coef-ql-nn
#| fig-cap: Coefficients, gain and phase shift functions for the Quadratic-Linear (QL) filter with $I/C=3.5$ and a nearest neighbour bandwidth.
img <- sprintf("img/filters_used/%s_nn.%s", tolower("QL"), "pdf")
knitr::include_graphics(img)
```

```{r, out.width="100%"}
#| label: fig-graphs-coef-cq-nn
#| fig-cap: Coefficients, gain and phase shift functions for the Cubic-Quadratic (CQ) filter with $I/C=3.5$ and a nearest neighbour bandwidth.
img <- sprintf("img/filters_used/%s_nn.%s", tolower("CQ"), "pdf")
knitr::include_graphics(img)
```

```{r, out.width="100%"}
#| label: fig-graphs-coef-daf-nn
#| fig-cap: Coefficients, gain and phase shift functions for the direct asymmetric filter (DAF) with a nearest neighbour bandwidth.
img <- sprintf("img/filters_used/%s_nn.%s", tolower("DAF"), "pdf")
knitr::include_graphics(img)
```

::: {#fig-nn-lp}
```{r, echo=FALSE, out.width="100%"}
knitr::include_graphics("img/simulations/phase_shift_simul_nn_lp.pdf")
```

Distribution of phase shift using nearest neighbour bandwidth.
:::

::: {#fig-nn-localparam}
```{r, echo=FALSE, out.width="100%"}
knitr::include_graphics("img/simulations/phase_shift_simul_nn_lp_localparam.pdf")
```

\justifying
*Note: nearest bandwidth moving average are used to estimate slope and concavity.*
*We thus do not distinguish local parametrisation and final local parametrisation which asumptions might not be plausible.*
*Indeed, using the same framework as for fixed bandwidth, the final local parametrisation for real time estimates would implie modelling a polynomial of degree 2 or 3 other 25 months (which doesn't seem plausible).*

Distribution of phase shift using nearest neighbour bandwidth for polynomial filters with local parametrisation.
:::

\newpage

# Quarterly data

::: {#fig-graphs-quarterly}
```{r, echo=FALSE, out.width="100%"}
knitr::include_graphics("img/simulations/phase_shift_simul_q.pdf")
```

Distribution of phase shift on quarterly data.
:::

::: {#tbl-simulrev-q}
```{r}
#| echo: false
rev_table <- readRDS("data/simulations_revisions_q.RDS")
pack_row_index = c("$MAE_{fe}(q) = \\mathbb E\\\\left[\\\\left|(TC_{t|t+q} -  TC_{t|last})/TC_{t|last}\\\\right|\\\\right]$" = nrow(rev_table) / 2,
                   "$MAE_{ce}(q)=\\mathbb E\\\\left[
\\\\left|(TC_{t|t+q} - TC_{t|t+q+1})/TC_{t|t+q+1}\\\\right|
\\\\right]$" = nrow(rev_table) / 2)
rev_table  %>%
  kable(format.args = list(digits = 2,
                           decimal.mark = "."),
        align = "c", booktabs = T, row.names = FALSE, 
        escape = FALSE,format = "latex") %>%  
  # kable_styling(latex_options=c(#"striped",  
  #   "hold_position")) %>% 
  pack_rows(index = pack_row_index, escape = FALSE) # %>%
    # add_footnote_kable(stop_centering = TRUE)
```

Average of the relative deviations of the revisions for the different filters on simulated quarterly series with medium variability.
:::